---
title: "Additional"
author: "Wu Jiayan"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **12  Visualising Models**

## **12.1 Introduction**

This section demonstrates visualising model diagnostic and model parameters by using parameters package.

## **12.2 Visual Analytics for Building Better Explanatory Models**

### **12.2.1 The case**

Toyota Corolla case study will be used. The purpose of study is to build a model to discover factors affecting prices of used-cars by taking into consideration a set of explanatory variables.

## **12.3 Preparation**

## **12.4 Installing and loading the required libraries**

```{r} pacman::p_load(readxl, SmartEDA, tidyverse,                ggstatsplot, easystats, tidymodels)}
```

### **12.4.1 Importing Excel file: readxl methods**

In the code chunk below, [*read_xls()*](https://readxl.tidyverse.org/reference/read_excel.html) of [**readxl**](https://readxl.tidyverse.org/) package is used to import the data worksheet of `ToyotaCorolla.xls` workbook into R.

```{r} car_resale <- read_xls("data/ToyotaCorolla.xls",                         "data") car_resale}
```

### **12.4.2 Visualising modelling variables**

```{r} ExpCatStat(car_resale,            Target = "Price",            result = "Stat")}
```

### 
