---
title: "Hands-on Exercise 4c: Visualising Uncertainty"
author: "Wu Jiayan"
date: "May 3, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **7  Visualising Uncertainty**

## **7.1 Learning Outcome**

-   to plot statistics error bars by using ggplot2,

-   to plot interactive error bars by combining ggplot2, plotly and DT,

-   to create advanced by using ggdist, and

-   to create hypothetical outcome plots (HOPs) by using ungeviz package.

## **7.2 Preparation**

### **7.2.1 Installing and loading the packages**

The following R packages will be used:

-   tidyverse, a family of R packages for data science process,

-   plotly for creating interactive plot,

-   gganimate for creating animation plot,

-   DT for displaying interactive html table,

-   crosstalk for for implementing cross-widget interactions (currently, linked brushing and filtering), and

-   ggdist for visualising distribution and uncertainty.

```{r}
pacman::p_load(plotly, crosstalk, DT, 
               ggdist, ggridges, colorspace,
               gganimate, tidyverse)
```

### **7.2.2 Data import**

```{r}
exam <- read_csv("data/Exam_data.csv")
```

## **7.3 Visualizing the uncertainty of point estimates: ggplot2 methods**

A point estimate is a single number, such as a mean. Uncertainty, on the other hand, is expressed as standard error, confidence interval, or credible interval.

This section demonstrates how to plot error bars of maths scores by race by using data provided in *exam* tibble data frame.

Firstly, code chunk below will be used to derive the necessary summary statistics.

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))
```

Key knowledge:

-   `group_by()` of **dplyr** package is used to group the observation by RACE,

-   `summarise()` is used to compute the count of observations, mean, standard deviation

-   `mutate()` is used to derive standard error of Maths by RACE, and

-   the output is save as a tibble data table called *my_sum*.

Next, the code chunk below will be used to display *my_sum* tibble data frame in an html table format.

```{r}
knitr::kable(head(my_sum), format = 'html')
```

### **7.3.1 Plotting standard error bars of point estimates**

Following code chunk plot the standard error bars of mean maths score by race:

::: panel-tabset
#### The Plot

```{r}
#| echo: false
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean maths score by race")
```

#### The Code

```{r}
#| eval: false
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean maths score by race")
```
:::

Key knowledge:

-   The error bars are computed by using the formula mean+/-se.

-   For `geom_point()`, it is important to indicate *stat=“identity”*. The code aims to plot pre-calculated mean values (y = mean) from the my_sum data frame.

### **7.3.2 Plotting confidence interval of point estimates**

Follow code chunk plots the confidence intervals of mean maths score by race.

::: panel-tabset
#### The Plot

```{r}
#| echo: false
ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE, -mean), 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  labs(x = "Maths score",
       title = "95% confidence interval of mean maths score by race")
```

#### The Code

```{r}
#| eval: false
ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE, -mean), 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  labs(x = "Maths score",
       title = "95% confidence interval of mean maths score by race")
```
:::

Key knowledge:

-   The confidence intervals are computed by using the formula mean+/-1.96\*se.

-   The error bars is sorted by using the average maths scores.

-   `labs()` argument of ggplot2 is used to change the x-axis label.

### **7.3.3 Visualizing the uncertainty of point estimates with interactive error bars**

Following code chunk plots interactive error bars for the 99% confidence interval of mean maths score by race.

::: panel-tabset
#### The Plot

```{r}
#| echo: false
shared_df = SharedData$new(my_sum)

bscols(widths = c(4,8),
       ggplotly((ggplot(shared_df) +
                   geom_errorbar(aes(
                     x=reorder(RACE, -mean),
                     ymin=mean-2.58*se, 
                     ymax=mean+2.58*se), 
                     width=0.2, 
                     colour="black", 
                     alpha=0.9, 
                     size=0.5) +
                   geom_point(aes(
                     x=RACE, 
                     y=mean, 
                     text = paste("Race:", `RACE`, 
                                  "<br>N:", `n`,
                                  "<br>Avg. Scores:", round(mean, digits = 2),
                                  "<br>95% CI:[", 
                                  round((mean-2.58*se), digits = 2), ",",
                                  round((mean+2.58*se), digits = 2),"]")),
                     stat="identity", 
                     color="red", 
                     size = 1.5, 
                     alpha=1) + 
                   xlab("Race") + 
                   ylab("Average Scores") + 
                   theme_minimal() + 
                   theme(axis.text.x = element_text(
                     angle = 45, vjust = 0.5, hjust=1)) +
                   ggtitle("99% Confidence interval of average /<br>maths scores by race")), 
                tooltip = "text"), 
       DT::datatable(shared_df, 
                     rownames = FALSE, 
                     class="compact", 
                     width="100%", 
                     options = list(pageLength = 10,
                                    scrollX=T), 
                     colnames = c("No. of pupils", 
                                  "Avg Scores",
                                  "Std Dev",
                                  "Std Error")) %>%
         formatRound(columns=c('mean', 'sd', 'se'),
                     digits=2))
```

#### The Code

```{r}
#| eval: false
shared_df = SharedData$new(my_sum)

bscols(widths = c(4,8),
       ggplotly((ggplot(shared_df) +
                   geom_errorbar(aes(
                     x=reorder(RACE, -mean),
                     ymin=mean-2.58*se, 
                     ymax=mean+2.58*se), 
                     width=0.2, 
                     colour="black", 
                     alpha=0.9, 
                     size=0.5) +
                   geom_point(aes(
                     x=RACE, 
                     y=mean, 
                     text = paste("Race:", `RACE`, 
                                  "<br>N:", `n`,
                                  "<br>Avg. Scores:", round(mean, digits = 2),
                                  "<br>95% CI:[", 
                                  round((mean-2.58*se), digits = 2), ",",
                                  round((mean+2.58*se), digits = 2),"]")),
                     stat="identity", 
                     color="red", 
                     size = 1.5, 
                     alpha=1) + 
                   xlab("Race") + 
                   ylab("Average Scores") + 
                   theme_minimal() + 
                   theme(axis.text.x = element_text(
                     angle = 45, vjust = 0.5, hjust=1)) +
                   ggtitle("99% Confidence interval of average /<br>maths scores by race")), 
                tooltip = "text"), 
       DT::datatable(shared_df, 
                     rownames = FALSE, 
                     class="compact", 
                     width="100%", 
                     options = list(pageLength = 10,
                                    scrollX=T), 
                     colnames = c("No. of pupils", 
                                  "Avg Scores",
                                  "Std Dev",
                                  "Std Error")) %>%
         formatRound(columns=c('mean', 'sd', 'se'),
                     digits=2))
```
:::

## **7.4 Visualising Uncertainty: ggdist package**

-   [**ggdist**](https://mjskay.github.io/ggdist/index.html) is an R package that provides a flexible set of ggplot2 geoms and stats designed especially for visualising distributions and uncertainty.

-   It is designed for both frequentist and Bayesian uncertainty visualization, taking the view that uncertainty visualization can be unified through the perspective of distribution visualization:

    -   for frequentist models, one visualises confidence distributions or bootstrap distributions;

    -   for Bayesian models, one visualises probability distributions (see the tidybayes package, which builds on top of ggdist).

### **7.4.1 Visualizing the uncertainty of point estimates: ggdist methods**

In the code chunk below, [`stat_pointinterval()`](https://mjskay.github.io/ggdist/reference/stat_pointinterval.html) of **ggdist** is used to build a visual for displaying distribution of maths scores by race.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

This function comes with many arguments, for example, in the code chunk below the following arguments are used:

-   .width = 0.95

-   .point = median

-   .interval = qi

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
                     point_interval = median_qi) +
  labs(
    title = "Visualising confidence intervals of median math score",
    subtitle = "Median Point + Multiple-interval plot")
```

### **7.4.2 Visualizing the uncertainty of point estimates: ggdist methods**

Code chunk below shows 95% and 99% confidence intervals:

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = c(0.95, 0.99),
                     point_interval = median_qi) +
  labs(
    title = "Visualising confidence intervals of median math score",
    subtitle = "Median Point + Multiple-interval plot")
```

### **7.4.3 Visualizing the uncertainty of point estimates: ggdist methods**

In the code chunk below, [`stat_gradientinterval()`](https://mjskay.github.io/ggdist/reference/stat_gradientinterval.html) of **ggdist** is used to build a visual for displaying distribution of maths scores by race.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")
```

## **7.5 Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)**

### **7.5.1 Installing ungeviz package**

Run installation if have not installed ungeviz package before:

```{r}
devtools::install_github("wilkelab/ungeviz")
```

### **7.5.2 Launch the application in R**

```{r}
library(ungeviz)
```

### **7.5.3 Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)**

The code chunk below will be used to build the HOPs.

```{r}
ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)
```
