---
title: "Hands-on Exercise 7b: Visualising and Analysing Text Data"
author: "Wu Jiayan"
date: "May 30, 2026"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
freeze: true
fig-width: 16
fig-height: 18
out-width: "100%"
---

# **15  Visualising and Analysing Text Data with R: tidytext methods**

## **29.1 Learning Outcome**

Learning Objectives:

- understand tidytext framework for processing, analysing and visualising text data,

- write function for importing multiple files into R,

- combine multiple files into a single data frame,

- clean and wrangle text data by using tidyverse approach,

- visualise words with Word Cloud,

- compute term frequency–inverse document frequency (TF-IDF) using tidytext method, and

- visualising texts and terms relationship.

## **29.2 Getting Started**

### **29.2.1 Installing and launching R packages**

In this hands-on exercise, the following R packages for handling, processing, wrangling, analysing and visualising text data will be used:

- tidytext, tidyverse (mainly readr, purrr, stringr, ggplot2)

- widyr,

- wordcloud and ggwordcloud,

- textplot (required igraph, tidygraph and ggraph, )

- DT,

- lubridate and hms.

The code chunk:

```{r}
pacman::p_load(tidytext, widyr, wordcloud, DT, ggwordcloud, textplot, lubridate, hms,
               tidyverse, tidygraph, ggraph, igraph)
```

## **29.3 Importing Multiple Text Files from Multiple Folders**

### **29.3.1 Creating a folder list**

```{r}
news20 <- "data/20news/"
```

### **29.3.2 Define a function to read all files from a folder into a data frame**

```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder, 
                    full.names = TRUE)) %>%
    mutate(text = map(file, 
                      read_lines)) %>%
    transmute(id = basename(file), 
              text) %>%
    unnest(text)
}
```

## **29.4 Importing Multiple Text Files from Multiple Folders**

### **29.4.1 Reading in all the messages from the 20news folder**

```{r}
raw_text <- tibble(folder = 
                     dir(news20, 
                         full.names = TRUE)) %>%
  mutate(folder_out = map(folder, 
                          read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), 
            id, text)
write_rds(raw_text, "data/rds/news20.rds")
```
