---
title: "Hands-on Exercise 2"
author: "Wu Jiayan"
date: "April 20, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **2 Beyond ggplot2 Fundamentals**

## **2.1 Overview**

This chapter introduces several ggplot2 extensions for creating more elegant and effective statistical graphics. Key knowledge and skills:

-   control the placement of annotation on a graph by using functions provided in ggrepel package,

-   create professional publication quality figure by using functions provided in ggthemes and hrbrthemes packages,

-   plot composite figure by combining ggplot2 graphs using patchwork package.

## **2.2 Getting started**

### **2.2.1 Installing and loading the required libraries**

Beside tidyverse, four R packages will be used:

-   ggrepel: an R package provides geoms for ggplot2 to repel overlapping text labels.

-   ggthemes: an R package provides some extra themes, geoms, and scales for ‘ggplot2’.

-   hrbrthemes: an R package provides typography-centric themes and theme components for ggplot2.

-   patchwork: an R package for preparing composite figure created using ggplot2.

Following code chunk check if these packages have been installed and also will load them onto your working R environment.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

***Troubleshooting:***

*If hrbrthemes are not yet available for newer version of R, run the following (optional):*

```{r}
#| eval: false
if (!require("remotes")) install.packages("remotes")
remotes::install_github("hrbrmstr/hrbrthemes")
```

### **2.2.2 Importing data**

*Exam_data* will be used which consists of year end examination grades of a cohort of primary 3 students from a local school. It is in csv file format.

The code chunk below imports *exam_data.csv* into R environment by using *read_csv()* function of [**readr**](https://readr.tidyverse.org/) package. **readr** is one of the tidyverse package.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

There are a total of seven attributes in the exam_data data frame. Four of them are categorical data type and the other three are in continuous data type.

-   The categorical attributes are: ID, CLASS, GENDER and RACE.

-   The continuous attributes are: MATHS, ENGLISH and SCIENCE.

## **2.3 Beyond ggplot2 Annotation: ggrepel**

One of the challenge in plotting statistical graph is annotation, especially with large number of data points.

::: panel-tabset
## The Plot

```{r}
#| echo: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## The Code

```{r}
#| eval: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

ggrepel is an extension of ggplot2 package which provides geoms for ggplot2 to repel overlapping text, simply replace geom_text() by geom_text_repel() and geom_label() by geom_label_repel.

### **2.3.1 Working with ggrepel**

::: panel-tabset
## The Plot

```{r}
#| echo: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## The Code

```{r}
#| eval: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

## **2.4 Beyond ggplot2 Themes**

ggplot2 comes with eight built-in themes: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void().

Following code chunk demonstrates theme_light:

::: panel-tabset
## The Plot

```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_light() +
  ggtitle("Distribution of Maths scores") 
```

## The Code

```{r}
#| eval: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_light() +
  ggtitle("Distribution of Maths scores") 
```
:::

### **2.4.1 Working with ggtheme package**

[**ggthemes**](https://cran.r-project.org/web/packages/ggthemes/index.html) provides [‘ggplot2’ themes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/) that replicate the look of plots by Edward Tufte, Stephen Few, [Fivethirtyeight](https://fivethirtyeight.com/), [The Economist](https://www.economist.com/graphic-detail), ‘Stata’, ‘Excel’, and [The Wall Street Journal](https://www.pinterest.com/wsjgraphics/wsj-graphics/), among others.

Following code chunk uses Economist theme:

::: panel-tabset
## The Plot

```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

## The Code

```{r}
#| eval: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
:::

### **2.4.2 Working with hrbthems package**

[**hrbrthemes**](https://cran.r-project.org/web/packages/hrbrthemes/index.html) package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

Following code chunk uses theme_ipsum

::: panel-tabset
## The Plot

```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

## The Code

```{r}
#| eval: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
:::

The second goal centers around productivity for a production workflow. In fact, this “production workflow” is the context for where the elements of hrbrthemes should be used. Check [this link](https://hrbrmstr.github.io/hrbrthemes/articles/why_hrbrthemes.html) for more information.

Following code chunk demonstrates the following:

-   `axis_title_size` argument is used to increase the font size of the axis title to 18,

-   `base_size` argument is used to increase the default axis label to 15, and

-   `grid` argument is used to remove the x-axis grid lines.

::: panel-tabset
## The Plot

```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

## The Code

```{r}
#| eval: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```
:::

***Troubleshooting:***

*If there is warning "font family not found in Windows font database", run the following code (optional):*

```{r}
#| eval: false
library(extrafont)
loadfonts(device = "win")
```

## **2.5 Beyond Single Graph**

In some cases, multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graphs.

Following code chunks create three statistical graphics

**1. Distribution of Match scores**

::: panel-tabset
## The Plot

```{r}
#| echo: false
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
p1
```

## The Code

```{r}
#| eval: false
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
```
:::

**2. Distribution of English score**

::: panel-tabset
## The Plot

```{r}
#| echo: false
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
p2
```

## The Code

```{r}
#| eval: false
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```
:::

**3. Scatterplot for English score versus Maths score**

::: panel-tabset
## The Plot

```{r}
#| echo: false
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
p3
```

## The Code

```{r}
#| eval: false
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

### **2.5.1 Creating Composite Graphics: pathwork methods**

There are several ggplot2 extension’s functions support the needs to prepare composite figure by combining several graphs such as [`grid.arrange()`](https://cran.r-project.org/web/packages/gridExtra/vignettes/arrangeGrob.html) of **gridExtra** package and [`plot_grid()`](https://wilkelab.org/cowplot/reference/plot_grid.html) of [**cowplot**](https://wilkelab.org/cowplot/index.html) package. Patchwork extension is specially designed for combining separate ggplot2 graphs into a single figure. This package has a very simple syntax where we can create layouts super easily. Here’s the general syntax that combines:

-   Two-Column Layout using the Plus Sign +.

-   Parenthesis () to create a subplot group.

-   Two-Row Layout using the Division Sign `/`

### **2.5.2 Combining two ggplot2 graphs**

Figure below shows a composite of two histograms created using patchwork, it uses very simple syntax to create the plot.

::: panel-tabset
## The Plot

```{r}
#| echo: false
#| fig-width: 12
p1 + p2
```

## The Code

```{r}
#| eval: false
p1 + p2
```
:::

### **2.5.3 Combining three ggplot2 graphs**

We can plot more complex composite by using appropriate operators. For example, the composite figure below is plotted by using:

-   “/” operator to stack two ggplot2 graphs,

-   “\|” operator to place the plots beside each other,

-   “()” operator the define the sequence of the plotting.

::: panel-tabset
## The Plot

```{r}
#| echo: false
#| out-width: "100%"
#| fig-width: 12
(p1 / p2) | p3
```

## The Code

```{r}
#| eval: false
#| out-width: "100%"
#| fig-width: 12
(p1 / p2) | p3
```
:::

### **2.5.4 Creating a composite figure with tag**

In order to identify subplots in text, patchwork also provides auto-tagging capabilities as shown in the figure below.

::: panel-tabset
## The Plot

```{r}
#| echo: false
#| out-width: "100%"
#| fig-width: 12
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = '1')
```

## The Code

```{r}
#| eval: false
#| out-width: "100%"
#| fig-width: 12
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = '1')
```
:::

Other tagging options:

|                |                                       |
|----------------|---------------------------------------|
| **Level Code** | **Result**                            |
| `'a'`          | Lowercase letters (a, b, c)           |
| `'A'`          | Uppercase letters (A, B, C)           |
| `'1'`          | Numbers (1, 2, 3)                     |
| `'i'`          | Lowercase Roman numerals (i, ii, iii) |
| `'I'`          | Uppercase Roman numerals (I, II, III) |

### **2.5.5 Creating figure with insert**

The [inset_element()](https://patchwork.data-imaginist.com/reference/inset_element.html) of patchwork can place one or several plots or graphic elements freely on top or below another plot, as demonstrated in the following code chunk:

::: panel-tabset
## The Plot

```{r}
#| echo: false
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

## The Code

```{r}
#| eval: false
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```
:::

### **2.5.6 Creating a composite figure by using patchwork and ggtheme**

Code chunk combines patchwork and theme_economist() of ggthemes package.

::: panel-tabset
## The Plot

```{r}
#| echo: false
#| out-width: "100%"
#| fig-width: 12
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

## The Code

```{r}
#| eval: false
#| out-width: "100%"
#| fig-width: 12
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
:::

## **2.6 Reference**

-   [Patchwork R package goes nerd viral](https://www.littlemissdata.com/blog/patchwork)

-   [ggrepel](https://ggrepel.slowkow.com/)

-   [ggthemes](https://ggplot2.tidyverse.org/reference/ggtheme.html)

-   [hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/)

-   [ggplot tips: Arranging plots](https://albert-rapp.de/post/2021-10-28-extend-plot-variety/)

-   [ggplot2 Theme Elements Demonstration](https://henrywang.nl/ggplot2-theme-elements-demonstration/)

-   [ggplot2 Theme Elements Reference Sheet](https://isabella-b.com/blog/ggplot2-theme-elements-reference/)
