---
title: "Hands-on Exercise 5d: Visual Multivariate Analysis with Parallel Coordinates Plot"
author: "Wu Jiayan"
date: "May 10, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
  fig-width: 14
  fig-height: 10
  out-width: "100%"
---

# **12  Visual Multivariate Analysis with Parallel Coordinates Plot**

## **12.1 Overview**

Parallel coordinates plot is a data visualisation specially designed for visualising and analysing multivariate, numerical data. It is ideal for comparing multiple variables together and seeing the relationships between them. For example, the variables contribute to Happiness Index. Parallel coordinates was invented by [Alfred Inselberg](http://www.math.tau.ac.il/~aiisreal/) in the 1970s as a way to visualize high-dimensional data. This data visualisation technique is more often found in academic and scientific communities than in business and consumer data visualizations. As pointed out by [Stephen Few](https://www.perceptualedge.com/articles/b-eye/parallel_coordinates.pdf)(2006), “This certainly isn’t a chart that you would present to the board of directors or place on your Web site for the general public. In fact, the strength of parallel coordinates isn’t in their ability to communicate some truth in the data to others, but rather in their ability to bring meaningful multivariate patterns and comparisons to light when used interactively for analysis.” For example, parallel coordinates plot can be used to characterise clusters detected during customer segmentation.

This hands-on exercise practices on:

-   plotting statistic parallel coordinates plots by using [`ggparcoord()`](https://ggobi.github.io/ggally/reference/ggparcoord.html) of [**GGally**](https://r4va.netlify.app/chap15) package, and

-   plotting interactive parallel coordinates plots by using [**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/) package.

## **12.2 Installing and Launching R Packages**

For this exercise, the **GGally**, **parallelPlot** and **tidyverse** packages will be used.

The code chunks below are used to install and load the packages in R.

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

## **12.3 Data Preparation**

In the code chunk below, `read_csv()` of **readr** package is used to import *WHData-2018.csv* into R and save it into a tibble data frame object called *wh*.

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

### **12.4.1 Plotting a simple parallel coordinates**

Code chunk below shows a typical syntax used to plot a basic static parallel coordinates plot by using `ggparcoord()`.

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

Only two argument namely `data` and `columns` is used. `Data` argument is used to map the data object (i.e. *wh*) and `columns` is used to select the columns for preparing the parallel coordinates plot.

### **12.4.2 Plotting a parallel coordinates with boxplot**

The basic parallel coordinates failed to reveal any meaning understanding of the World Happiness measures. Following sections makeover the plot by using a collection of arguments provided by `ggparcoord()`.

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

### **12.4.3 Parallel coordinates with facet**

Since `ggparcoord()` is developed by extending **ggplot2** package, we can combine some of the ggplot2 function when plotting a parallel coordinates plot.

In the code chunk below, `facet_wrap()` of **ggplot2** is used to plot 10 small multiple parallel coordinates plots. Each plot represent one geographical region such as East Asia.

```{r}
ggparcoord(data = wh, columns = c(7:12), groupColumn = 2, scale = "uniminmax", alphaLines = 0.2, boxplot = TRUE, title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") + facet_wrap(~ Region)
```

One of the aesthetic defect of the current design is that some of the variable names overlap on x-axis.

### **12.4.4 Rotating x-axis text label**

To make the x-axis text label easy to read, let us rotate the labels by 30 degrees. We can rotate axis text labels using `theme()` function in ggplot2 as shown in the code chunk below

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```

Thing to learn from the code chunk above:

-   To rotate x-axis text labels, we use `axis.text.x` as argument to `theme()` function. And we specify `element_text(angle = 30)` to rotate the x-axis text by an angle 30 degree.

### **12.4.5 Adjusting the rotated x-axis text label**

Rotating x-axis text labels to 30 degrees makes the label overlap with the plot and we can avoid this by adjusting the text location using `hjust` argument to theme’s text element with `element_text()`. We use `axis.text.x` as we want to change the look of x-axis text.

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

## **12.5 Plotting Interactive Parallel Coordinates Plot: parallelPlot methods**

[**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/) is an R package specially designed to plot a parallel coordinates plot by using ‘htmlwidgets’ package and [d3.js](https://d3js.org/). This section demonstrates how to use functions provided in parallelPlot package to build interactive parallel coordinates plot.

### **12.5.1 The basic plot**

The code chunk below plot an interactive parallel coordinates plot by using `parallelPlot()`.

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)
```

Issue: some of the axis labels are too long.

### **12.5.2 Rotate axis label**

In the code chunk below, `rotateTitle` argument is used to avoid overlapping axis labels.

```{r}
parallelPlot(wh,
             rotateTitle = TRUE)
```

One of the useful interactive feature of parallelPlot is we can click on a variable of interest, for example Happiness score, the monotonous blue colour (default) will change a blues with different intensity colour scheme will be used.

### **12.5.3 Changing the colour scheme**

Following code chunk changes the default blue colour scheme by using `continousCS` argument.

```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

### **12.5.4 Parallel coordinates plot with histogram**

In the code chunk below, `histoVisibility` argument is used to plot histogram along the axis of each variables.

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```

## **12.6 References**

-   [*ggparcoord()*](http://ggobi.github.io/ggally/reference/ggparcoord.html) of [**GGally**](http://ggobi.github.io/ggally/index.html) package

-   [**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/parallelPlot.pdf)
