---
title: "Visual Analytics for Motor Insurance Portfolio Discovery"
author: "Wu Jiayan"
date: "May 10, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    toc-location: right
    number-sections: true
    code-fold: false
    smooth-scroll: true
    fig-align: center
    embed-resources: true
    html-math-method: katex
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
  fig-width: 16
  fig-height: 8
  out-width: "100%"
---

# Introduction

## Background and Objectives

This report documents the complete visual analytics investigation of a real-world motor vehicle insurance portfolio from a Spanish non-life insurer. The dataset, sourced from [Mendeley Data](https://data.mendeley.com/datasets/sw4jmdb2sm/1), contains policy-year records spanning 2022–2024 with information on policyholder demographics, vehicle characteristics, coverage types, premiums, and claims outcomes.

**Investigation objectives:**

1.  Apply **univariate visual analytics** to characterise the distributions and composition of numerical and categorical variables.
2.  Apply **bivariate visual analytics** to investigate and statistically validate factors influencing portfolio **profitability** (loss ratio, profit per policy) and **volatility** (claim frequency, claim severity).

## Dataset Description

The dataset is organised into four variable groups as defined in the official data cookbook:

**Policy and Insured Characteristics** — variables describing the contract itself:

| Variable | Description |
|----|----|
| `insured_id` | Unique sequential ID per policy number; repeated policies share the same ID |
| `year` | Calendar year of the record |
| `policy_type` | Coverage structure: TP (basic liability), TPG (liability + glass), CC (liability + 2+ extra covers), COMP_N (comprehensive without excess), COMP_E (comprehensive with excess) |
| `policy_status` | Active (A) or Cancelled (C) at time of analysis |
| `business_type` | New Business (NB) or Portfolio renewal (P) |
| `payment_frequency` | Annual (A), Semi-annual (S), or Quarterly (Q) billing |
| `bonus_score` | Past claims experience: Good/G (favourable), Neutral/N, Bad/B (poor history) |

**Driver and Vehicle Characteristics** — risk-rating variables:

| Variable | Description |
|----|----|
| `driver_age` | Age of the main driver in years |
| `vehicle_age` | Age of the insured vehicle in years |
| `age_driving_licence` | Calendar year the licence was issued *(not* years of experience held) |
| `fuel_type` | Gasoline (G) or Diesel (D) only — other fuel types excluded |
| `vehicle_value` | Declared insured value; some entries missing |
| `seats` | Number of seats in the vehicle |
| `power_to_weight_ratio` | Approximate kg per horsepower — proxy for performance |
| `vehicle_brand` | Brand as reported in policy data |
| `municipality_type` | Inland (I), Coastal (C), or Islands (IS) |
| `circulation_area` | Predominant driving environment: Urban (U) or Rural (R) |

**Premiums** — eight coverage-level premium components that sum to `total_premium`: liability, property damage, theft, fire, glass, legal protection, occupants, plus the total.

**Exposure, Claims and Incurred** — outcome variables: `total_exposure` (policy years on risk, 0–1 for partial years, may exceed 1 for multi-year contracts), claim counts per coverage line, and incurred costs (paid amounts plus outstanding reserves) per coverage line.

# Getting Started

Following R packages will be used:

-   tidyverse: Core data manipulation (dplyr, tidyr, readr, forcats, stringr)

-   ggplot2: Grammar of graphics visualisation engine

-   scales: Axis formatting helpers (percent, comma, dollar)

-   patchwork: Multi-panel plot composition

-   ggridges: Ridge / joy plots for distributional comparison

-   ggcorrplot: Correlation matrix heat maps

-   ggrepel: Non-overlapping text labels

-   knitr: Table rendering in Quarto

-   kableExtra: HTML table styling

-   ggiraph: interactive ggplot2 geoms (tooltip, hover, click)

-   plotly: interactive charts via ggplotly() or plot_ly()

-   DT: interactive JavaScript DataTables

```{r}
#| code-fold: true
#| code-summary: "Show R Code"
pacman::p_load(ggcorrplot, ggplot2, tidyverse, scales, patchwork, ggridges, ggrepel,
               knitr, kableExtra,
               ggiraph,  # interactive ggplot2 geoms (tooltip, hover) — Chap 03 §3.4-3.6
               plotly,   # interactive charts via ggplotly() and plot_ly() — Chap 03 §3.7
               DT)       # interactive DataTables for R — Chap 03 §3.8
```

Following code chunk defines visual style and color palettes to ensure all charts in the report look consistent. 

```{r setup}
#| label: setup
#| include: true
#| code-fold: true
#| code-summary: "Show R Code"


# ── Global ggplot2 theme ─────────────────────────────────────────────────────
theme_ins <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title    = element_text(face = "bold", size = base_size + 1),
      plot.subtitle = element_text(colour = "grey50", size = base_size - 1),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom"
    )
}

theme_set(theme_ins())

# ── Consistent colour palette ────────────────────────────────────────────────
pal_main  <- c("#1565C0","#1E88E5","#42A5F5","#90CAF9","#E3F2FD")
pal_div   <- c("#D32F2F","#EF5350","#EF9A9A","#BBDEFB","#42A5F5","#1565C0")
pal_cat5  <- c("#1565C0","#FF9800","#4CAF50","#F44336","#9C27B0")
pal_bonus <- c("Good (Favourable History)" = "#4CAF50", Neutral = "#FF9800", "Bad (Poor History)" = "#F44336")
pal_area  <- c(Urban = "#FF9800", Rural = "#1E88E5")
pal_fuel  <- c(Gasoline = "#FF9800", Diesel = "#1E88E5")
```

# Data Preparation

## Loading Data

Following code chunk loads data into R environment as df_raw, read_delim() is used instead of read_csv() because the file uses semicolons as separators, show_col_types = FALSE suppresses the column-type printout.

```{r load-data}
#| label: load-data
#| code-fold: true
#| code-summary: "Show R Code"
#Reading the data into R environment
df_raw <- read_delim(
  "data/Dataset_of_motor_insurance_portfolio.csv",
  delim = ";",
  show_col_types = FALSE
)
```

## Variable Classification

Following code chunk builds a metadata table about the dataset itself, enriched with variable descriptions from the official data cookbook. Key functions:

-   `names(df_raw)` — extracts all column names

-   `sapply(df_raw, class)` — applies `class()` to every column to get its data type

-   `case_when()` — tidyverse's multi-condition `if-else`, here used to classify each variable into the four analytical groups defined in the cookbook: *Policy and insured characteristics*, *Driver and vehicle characteristics*, *Premiums*, and *Exposure, claims and incurred*

-   `str_detect(Variable, "premium")` — regex match to catch all premium-related columns without listing them individually

-   `colSums(is.na())` — counts missing values per column

::: panel-tabset
#### The Table

```{r tbl-var-types}
#| label: tbl-var-types
#| echo: false

var_types <- tibble(
  Variable = names(df_raw),
  Type = sapply(df_raw, class),
  Description = case_when(
    Variable == "insured_id"                  ~ "Unique sequential identifier per policy number",
    Variable == "year"                        ~ "Calendar year of the policy record",
    Variable == "policy_type"                 ~ "Coverage structure: TP / TPG / CC / COMP_N / COMP_E",
    Variable == "policy_status"               ~ "Active (A) or Cancelled (C) at time of analysis",
    Variable == "business_type"               ~ "New Business (NB) or Portfolio renewal (P)",
    Variable == "payment_frequency"           ~ "Billing frequency: Annual (A), Semi-annual (S), Quarterly (Q)",
    Variable == "bonus_score"                 ~ "Claims history indicator: Good (G), Neutral (N), Bad (B)",
    Variable == "driver_age"                  ~ "Age of the main driver in years",
    Variable == "vehicle_age"                 ~ "Age of the insured vehicle in years",
    Variable == "age_driving_licence"         ~ "Calendar year licence was issued (not years held)",
    Variable == "fuel_type"                   ~ "Vehicle fuel: Gasoline (G) or Diesel (D)",
    Variable == "vehicle_value"               ~ "Declared insured value of the vehicle (some missing)",
    Variable == "seats"                       ~ "Number of seats in the insured vehicle",
    Variable == "power_to_weight_ratio"       ~ "Approx. kg per horsepower — proxy for vehicle performance",
    Variable == "vehicle_brand"               ~ "Brand of the insured vehicle",
    Variable == "municipality_type"           ~ "Geographic classification: Inland (I), Coastal (C), Islands (IS)",
    Variable == "circulation_area"            ~ "Predominant environment: Urban (U) or Rural (R)",
    Variable == "total_premium"               ~ "Sum of all individual coverage premiums",
    Variable == "total_claims"                ~ "Total number of claims reported during exposure period",
    Variable == "total_incurred"              ~ "Total incurred cost: paid amounts plus outstanding reserves",
    Variable == "total_exposure"              ~ "Effective time-on-risk in policy years (0–1, may exceed 1)",
    Variable == "liability_exposure"          ~ "Exposure associated with liability coverage",
    str_detect(Variable, "^liability_p.*prem") ~ "Premium for liability property damage coverage",
    str_detect(Variable, "^liability_i.*prem") ~ "Premium for liability bodily injury coverage",
    str_detect(Variable, "liability_premium")  ~ "Premium charged for third-party liability coverage",
    str_detect(Variable, "property_damage_premium") ~ "Premium for property damage coverage",
    str_detect(Variable, "theft_premium")     ~ "Premium for theft coverage",
    str_detect(Variable, "fire_premium")      ~ "Premium for fire coverage",
    str_detect(Variable, "glass_premium")     ~ "Premium for glass coverage",
    str_detect(Variable, "legal_protection_premium") ~ "Premium for legal protection coverage",
    str_detect(Variable, "occupants_premium") ~ "Premium for personal-injury coverage of occupants",
    str_detect(Variable, "_claims$")          ~ paste0("Count of ", str_remove(Variable, "_claims"), " claims"),
    str_detect(Variable, "_incurred$")        ~ paste0("Incurred cost (paid + reserved) for ", str_remove(Variable, "_incurred"), " claims"),
    TRUE ~ Variable
  ),
  Group = case_when(
    Variable %in% c("insured_id","year","policy_type","policy_status",
                    "business_type","payment_frequency","bonus_score") ~ "Policy & Insured Characteristics",
    Variable %in% c("driver_age","vehicle_age","age_driving_licence","fuel_type",
                    "vehicle_value","seats","power_to_weight_ratio",
                    "vehicle_brand","municipality_type","circulation_area") ~ "Driver & Vehicle Characteristics",
    str_detect(Variable, "premium") ~ "Premiums",
    str_detect(Variable, "claims|incurred|exposure") ~ "Exposure, Claims & Incurred",
    TRUE ~ "Other"
  ),
  Missing_N = colSums(is.na(df_raw)),
  Missing_Pct = scales::percent(colSums(is.na(df_raw)) / nrow(df_raw), accuracy = 0.01)
)

library(htmltools)

# Build the table HTML
tbl_html <- kbl(var_types,
                caption = "Variable Classification and Missing Data Summary (Source: Mendeley Data Cookbook)",
                table.attr = 'id="var-types-tbl"') |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) |>
  column_spec(4, bold = TRUE) |>
  scroll_box(height = "400px")

# Strip kableExtra's hardcoded inline background-color / color from every cell
tbl_clean <- gsub(
  'style="([^"]*)(background-color:[^;]+;?)([^"]*)(color:[^;]+;?)([^"]*)"',
  'style="\\1\\3\\5"',
  as.character(tbl_html)
)
tbl_clean <- gsub(
  'style="([^"]*)(background-color:[^;]+;?)([^"]*)"',
  'style="\\1\\3"',
  tbl_clean
)

# Inject a <style> block scoped to #var-types-tbl so it adapts to light/dark
adaptive_css <- '
<style>
  /* Mid-tone #8fa8c4: contrast ≥3.0 against both white and black text */
  #var-types-tbl th {
    background-color: #4a6080 !important;
    color:            #ffffff !important;
    border-bottom:    2px solid #2d4a6a !important;
  }
  #var-types-tbl tr:nth-child(odd) td {
    background-color: #8fa8c4 !important;
    color:            #1a1a1a !important;
  }
  #var-types-tbl tr:nth-child(even) td {
    background-color: #c8d8e8 !important;
    color:            #1a1a1a !important;
  }
  #var-types-tbl tr:hover td {
    background-color: #7a9ab8 !important;
    color:            #1a1a1a !important;
  }
</style>
'

HTML(paste0(adaptive_css, tbl_clean))
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

var_types <- tibble(
  Variable = names(df_raw),
  Type = sapply(df_raw, class),
  Description = case_when(
    Variable == "insured_id"                  ~ "Unique sequential identifier per policy number",
    Variable == "year"                        ~ "Calendar year of the policy record",
    Variable == "policy_type"                 ~ "Coverage structure: TP / TPG / CC / COMP_N / COMP_E",
    Variable == "policy_status"               ~ "Active (A) or Cancelled (C) at time of analysis",
    Variable == "business_type"               ~ "New Business (NB) or Portfolio renewal (P)",
    Variable == "payment_frequency"           ~ "Billing frequency: Annual (A), Semi-annual (S), Quarterly (Q)",
    Variable == "bonus_score"                 ~ "Claims history indicator: Good (G), Neutral (N), Bad (B)",
    Variable == "driver_age"                  ~ "Age of the main driver in years",
    Variable == "vehicle_age"                 ~ "Age of the insured vehicle in years",
    Variable == "age_driving_licence"         ~ "Calendar year licence was issued (not years held)",
    Variable == "fuel_type"                   ~ "Vehicle fuel: Gasoline (G) or Diesel (D)",
    Variable == "vehicle_value"               ~ "Declared insured value of the vehicle (some missing)",
    Variable == "seats"                       ~ "Number of seats in the insured vehicle",
    Variable == "power_to_weight_ratio"       ~ "Approx. kg per horsepower — proxy for vehicle performance",
    Variable == "vehicle_brand"               ~ "Brand of the insured vehicle",
    Variable == "municipality_type"           ~ "Geographic classification: Inland (I), Coastal (C), Islands (IS)",
    Variable == "circulation_area"            ~ "Predominant environment: Urban (U) or Rural (R)",
    Variable == "total_premium"               ~ "Sum of all individual coverage premiums",
    Variable == "total_claims"                ~ "Total number of claims reported during exposure period",
    Variable == "total_incurred"              ~ "Total incurred cost: paid amounts plus outstanding reserves",
    Variable == "total_exposure"              ~ "Effective time-on-risk in policy years (0–1, may exceed 1)",
    Variable == "liability_exposure"          ~ "Exposure associated with liability coverage",
    str_detect(Variable, "^liability_p.*prem") ~ "Premium for liability property damage coverage",
    str_detect(Variable, "^liability_i.*prem") ~ "Premium for liability bodily injury coverage",
    str_detect(Variable, "liability_premium")  ~ "Premium charged for third-party liability coverage",
    str_detect(Variable, "property_damage_premium") ~ "Premium for property damage coverage",
    str_detect(Variable, "theft_premium")     ~ "Premium for theft coverage",
    str_detect(Variable, "fire_premium")      ~ "Premium for fire coverage",
    str_detect(Variable, "glass_premium")     ~ "Premium for glass coverage",
    str_detect(Variable, "legal_protection_premium") ~ "Premium for legal protection coverage",
    str_detect(Variable, "occupants_premium") ~ "Premium for personal-injury coverage of occupants",
    str_detect(Variable, "_claims$")          ~ paste0("Count of ", str_remove(Variable, "_claims"), " claims"),
    str_detect(Variable, "_incurred$")        ~ paste0("Incurred cost (paid + reserved) for ", str_remove(Variable, "_incurred"), " claims"),
    TRUE ~ Variable
  ),
  Group = case_when(
    Variable %in% c("insured_id","year","policy_type","policy_status",
                    "business_type","payment_frequency","bonus_score") ~ "Policy & Insured Characteristics",
    Variable %in% c("driver_age","vehicle_age","age_driving_licence","fuel_type",
                    "vehicle_value","seats","power_to_weight_ratio",
                    "vehicle_brand","municipality_type","circulation_area") ~ "Driver & Vehicle Characteristics",
    str_detect(Variable, "premium") ~ "Premiums",
    str_detect(Variable, "claims|incurred|exposure") ~ "Exposure, Claims & Incurred",
    TRUE ~ "Other"
  ),
  Missing_N = colSums(is.na(df_raw)),
  Missing_Pct = scales::percent(colSums(is.na(df_raw)) / nrow(df_raw), accuracy = 0.01)
)

kbl(var_types, caption = "Variable Classification and Missing Data Summary (Source: Mendeley Data Cookbook)") |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE) |>
  column_spec(4, bold = TRUE) |>
  scroll_box(height = "400px")
```
:::

## Feature Engineering

Derived variables are constructed to support the analytical objectives. The **loss ratio** (incurred ÷ premium) is the primary profitability metric in non-life insurance; **claim indicator** enables frequency modelling; **profit** facilitates absolute value analysis. Note that `total_exposure` — the effective time-on-risk expressed in policy years — ranges from 0 to 1 for partial-year policies and may exceed 1 for multi-year contracts. `total_incurred` captures both paid amounts and outstanding reserves, so it represents the full economic cost of claims, not just settled payments.

```{r feature-engineering}
#| label: feature-engineering
#| code-fold: true
#| code-summary: "Show R Code"

df <- df_raw |>
  mutate(
    # ── Temporal ──────────────────────────────────────────────────────────────
    year = factor(year),

    # ── Categorical recoding ──────────────────────────────────────────────────
    policy_type = factor(policy_type,
                         levels = c("TP","TPG","CC","COMP_N","COMP_E"),
                         labels = c("Third Party","Third Party+Glass",
                                    "Third Party+Multi-Cover","Comprehensive (No Excess)","Comprehensive (With Excess)")),
    policy_status = factor(policy_status, levels = c("A","C"),
                           labels = c("Active","Cancelled")),
    business_type = factor(business_type, levels = c("NB","P"),
                           labels = c("New Business","Portfolio")),
    payment_frequency = factor(payment_frequency,
                               levels = c("A","S","Q"),
                               labels = c("Annual","Semi-Annual","Quarterly")),
    bonus_score = factor(bonus_score, levels = c("G","N","B"),
                         labels = c("Good (Favourable History)","Neutral","Bad (Poor History)")),
    fuel_type = factor(fuel_type, levels = c("G","D"),
                       labels = c("Gasoline","Diesel")),
    municipality_type = factor(municipality_type,
                               levels = c("I","C","IS"),
                               labels = c("Inland","Coastal","Islands")),
    circulation_area = factor(circulation_area, levels = c("U","R"),
                              labels = c("Urban","Rural")),

    # ── Derived variables ─────────────────────────────────────────────────────
    has_claim   = total_claims > 0,
    loss_ratio  = if_else(total_premium > 0,
                          total_incurred / total_premium, NA_real_),
    profit      = total_premium - total_incurred,
    severity    = if_else(has_claim, total_incurred / total_claims, NA_real_),
    driver_age_group = cut(driver_age,
                           breaks = c(17,25,35,45,55,65,90),
                           labels = c("18-25","26-35","36-45","46-55","56-65","66+"),
                           right = TRUE),
    veh_value_q = ntile(vehicle_value, 5)
  )

# Summary of key derived variables
df |>
  select(total_premium, total_incurred, has_claim, loss_ratio, profit, severity) |>
  summary()
```

## Data Quality Assessment

::: panel-tabset
#### The Plot

```{r data-quality}
#| echo: false
#| label: data-quality

# Missing values by variable
missing_summary <- df_raw |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing") |>
  filter(Missing > 0) |>
  mutate(Pct = Missing / nrow(df_raw))

library(htmltools)

# Mid-tone table colours (readable by both black and white text)
dq_css <- '
<style>
  #dq-tbl th {
    background-color: #4a6080; color: #ffffff; border-bottom: 2px solid #2d4a6a;
  }
  #dq-tbl tr:nth-child(odd)  td { background-color: #8fa8c4; color: #1a1a1a; }
  #dq-tbl tr:nth-child(even) td { background-color: #c8d8e8; color: #1a1a1a; }
  #dq-tbl tr:hover td { background-color: #7a9ab8; color: #1a1a1a; }
</style>
'
dq_tbl <- kbl(missing_summary |>
      mutate(Pct = scales::percent(Pct, accuracy = 0.01)),
    caption = "Variables with Missing Values",
    col.names = c("Variable","Missing Count","Missing %"),
    table.attr = 'id="dq-tbl"') |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)

dq_clean <- gsub('style="([^"]*)(background-color:[^;]+;?)([^"]*)"',
                  'style="\1\3"', as.character(dq_tbl))
HTML(paste0(dq_css, dq_clean))
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

# Missing values by variable
missing_summary <- df_raw |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing") |>
  filter(Missing > 0) |>
  mutate(Pct = Missing / nrow(df_raw))

kbl(missing_summary |>
      mutate(Pct = scales::percent(Pct, accuracy = 0.01)),
    caption = "Variables with Missing Values",
    col.names = c("Variable","Missing Count","Missing %")) |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```
:::

Missing data is minimal: `fuel_type` (0.36%) and `vehicle_age` (\< 0.001%) have negligible missingness and are handled by dropping NA rows in fuel-specific analyses. No imputation is required given the tiny proportion.

------------------------------------------------------------------------

# Univariate Visual Analytics

Univariate analysis examines each variable in isolation to understand its distribution, central tendency, dispersion, and shape. This establishes a baseline understanding of the portfolio before investigating relationships between variables.

## Numerical Variables

### Justification of Technique

For continuous numerical variables, **histograms** reveal the full distributional shape including skewness, modality, and outlier presence. They are preferred over box plots for initial exploration when sample sizes are large (n \> 10,000) as they expose tail behaviour critical for insurance pricing. **Density plots** are overlaid where smooth distributional form aids interpretation.

### Total Annual Premium

::: panel-tabset
#### The Plot

```{r premium-dist}
#| echo: false
#| label: premium-dist
#| fig-cap: "Distribution of Total Annual Premium (truncated at 99.5th percentile)"

q995_prem <- quantile(df$total_premium, 0.995, na.rm = TRUE)

p1 <- df |>
  filter(total_premium > 0, total_premium <= q995_prem) |>
  ggplot(aes(total_premium)) +
  geom_histogram(aes(y = after_stat(density)), bins = 70,
                 fill = "#1E88E5", colour = NA, alpha = 0.85) +
  geom_density(colour = "#FF9800", linewidth = 0.9) +
  geom_vline(xintercept = median(df$total_premium[df$total_premium > 0], na.rm = TRUE),
             colour = "#F44336", linetype = "dashed", linewidth = 1) +
  annotate("text", x = median(df$total_premium[df$total_premium > 0], na.rm=TRUE) + 15,
           y = 0.002, label = paste0("Median: €",
             round(median(df$total_premium[df$total_premium>0],na.rm=TRUE))),
           colour = "#F44336", hjust = 0, size = 3.5) +
  scale_x_continuous(labels = scales::dollar_format(prefix = "€"),
                     name = "Total Annual Premium (€)") +
  scale_y_continuous(name = "Density") +
  labs(title = "Distribution of Total Annual Premium",
       subtitle = "Right-skewed; median ~€258; long upper tail from high-value covers")

p1
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

q995_prem <- quantile(df$total_premium, 0.995, na.rm = TRUE)

p1 <- df |>
  filter(total_premium > 0, total_premium <= q995_prem) |>
  ggplot(aes(total_premium)) +
  geom_histogram(aes(y = after_stat(density)), bins = 70,
                 fill = "#1E88E5", colour = NA, alpha = 0.85) +
  geom_density(colour = "#FF9800", linewidth = 0.9) +
  geom_vline(xintercept = median(df$total_premium[df$total_premium > 0], na.rm = TRUE),
             colour = "#F44336", linetype = "dashed", linewidth = 1) +
  annotate("text", x = median(df$total_premium[df$total_premium > 0], na.rm=TRUE) + 15,
           y = 0.002, label = paste0("Median: €",
             round(median(df$total_premium[df$total_premium>0],na.rm=TRUE))),
           colour = "#F44336", hjust = 0, size = 3.5) +
  scale_x_continuous(labels = scales::dollar_format(prefix = "€"),
                     name = "Total Annual Premium (€)") +
  scale_y_continuous(name = "Density") +
  labs(title = "Distribution of Total Annual Premium",
       subtitle = "Right-skewed; median ~€258; long upper tail from high-value covers")

p1
```
:::

The premium distribution is strongly **right-skewed** with a long upper tail. The median annual premium of approximately €258 indicates most policies are competitively priced, while the extended tail reflects high-value comprehensive covers.

### Total Incurred Claims

The incurred claims distribution is restricted to policyholders who made at least one claim (`total_incurred > 0`) and truncated at the 99.5th percentile to remove extreme outliers while preserving the shape of the loss distribution. Key functions:

-   `geom_histogram(aes(y = after_stat(density)))` — rescales the histogram to density so it shares the same y-axis as the overlaid `geom_density()` curve

-   `filter(total_incurred > 0)` — excludes the mass of zero-claim policies, isolating the severity distribution among claimants only

::: panel-tabset
#### The Plot

```{r incurred-dist}
#| echo: false
#| label: incurred-dist
#| fig-cap: "Distribution of Incurred Losses — claimants only (truncated at 99.5th percentile)"

q995_inc <- quantile(df$total_incurred[df$total_incurred > 0], 0.995, na.rm = TRUE)

p2 <- df |>
  filter(total_incurred > 0, total_incurred <= q995_inc) |>
  ggplot(aes(total_incurred)) +
  geom_histogram(aes(y = after_stat(density)), bins = 70,
                 fill = "#F44336", colour = NA, alpha = 0.8) +
  geom_density(colour = "#FF9800", linewidth = 0.9) +
  scale_x_continuous(labels = scales::dollar_format(prefix = "€")) +
  labs(title = "Distribution of Total Incurred Claims (Claimants Only)",
       subtitle = "Extreme right skew; heavy upper tail; consistent with Pareto/log-normal loss models",
       x = "Incurred Claims (€)", y = "Density")

p2
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

q995_inc <- quantile(df$total_incurred[df$total_incurred > 0], 0.995, na.rm = TRUE)

p2 <- df |>
  filter(total_incurred > 0, total_incurred <= q995_inc) |>
  ggplot(aes(total_incurred)) +
  geom_histogram(aes(y = after_stat(density)), bins = 70,
                 fill = "#F44336", colour = NA, alpha = 0.8) +
  geom_density(colour = "#FF9800", linewidth = 0.9) +
  scale_x_continuous(labels = scales::dollar_format(prefix = "€")) +
  labs(title = "Distribution of Total Incurred Claims (Claimants Only)",
       subtitle = "Extreme right skew; heavy upper tail; consistent with Pareto/log-normal loss models",
       x = "Incurred Claims (€)", y = "Density")

p2
```
:::

Claims incurred among policyholders who made at least one claim exhibit **extreme right skewness**, consistent with actuarial loss models (log-normal, gamma, or Pareto distributions). The heavy upper tail represents catastrophic losses and has significant reserve implications.

### Driver Age and Vehicle Age

::: panel-tabset
#### The Plot

```{r age-distributions}
#| echo: false
#| label: age-distributions
#| fig-cap: "Distributions of driver age (left) and vehicle age (right)"

p_dage <- df |>
  ggplot(aes(driver_age)) +
  geom_histogram(bins = 50, fill = "#42A5F5", colour = NA, alpha = 0.85) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Driver Age Distribution",
       subtitle = "Approximately normal; mean ~49 years; mode at age 40",
       x = "Driver Age (years)", y = "Count")

p_vage <- df |>
  filter(!is.na(vehicle_age)) |>
  ggplot(aes(vehicle_age)) +
  geom_histogram(bins = 50, fill = "#4CAF50", colour = NA, alpha = 0.85) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Vehicle Age Distribution",
       subtitle = "Right-skewed; highest concentration of vehicles aged 18–22 years",
       x = "Vehicle Age (years)", y = "Count")

p_dage + p_vage
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

p_dage <- df |>
  ggplot(aes(driver_age)) +
  geom_histogram(bins = 50, fill = "#42A5F5", colour = NA, alpha = 0.85) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Driver Age Distribution",
       subtitle = "Approximately normal; mean ~49 years; mode at age 40",
       x = "Driver Age (years)", y = "Count")

p_vage <- df |>
  filter(!is.na(vehicle_age)) |>
  ggplot(aes(vehicle_age)) +
  geom_histogram(bins = 50, fill = "#4CAF50", colour = NA, alpha = 0.85) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Vehicle Age Distribution",
       subtitle = "Right-skewed; highest concentration of vehicles aged 18–22 years",
       x = "Vehicle Age (years)", y = "Count")

p_dage + p_vage
```
:::

Driver age is approximately normally distributed with a mean of 49 years and mode at age 40, reflecting a mature policyholder base. Vehicle age shows a right-skewed distribution with the highest concentration of vehicles aged 18–22 years, suggesting the portfolio is dominated by older vehicles rather than new registrations. Note that `age_driving_licence` records the **calendar year** the licence was issued — not the number of years the licence has been held — and should be transformed (e.g., current year minus licence year) before use as an experience proxy in any model.

### Numerical Summary Statistics

::: panel-tabset
#### The Plot

```{r numerical-summary}
#| echo: false
#| label: numerical-summary

df |>
  select(driver_age, vehicle_age, vehicle_value, power_to_weight_ratio,
         total_premium, total_incurred, total_claims, total_exposure) |>
  summarise(across(everything(),
                   list(
                     Min    = ~min(., na.rm=TRUE),
                     Q1     = ~quantile(., 0.25, na.rm=TRUE),
                     Median = ~median(., na.rm=TRUE),
                     Mean   = ~mean(., na.rm=TRUE),
                     Q3     = ~quantile(., 0.75, na.rm=TRUE),
                     Max    = ~max(., na.rm=TRUE)
                   ))) |>
  pivot_longer(everything(), names_sep = "_(?=[^_]+$)", names_to = c("Variable","Stat")) |>
  pivot_wider(names_from = Stat, values_from = value) |>
  mutate(across(where(is.numeric), ~round(., 2))) |>
  # DT::datatable (Chap 03 §3.8): sortable, searchable, paginated table.
  # initComplete applies mid-tone row colours (#8fa8c4) directly via JS,
  # bypassing DataTables' own CSS so colours work regardless of theme.
  DT::datatable(
    caption  = "Descriptive Statistics for Numerical Variables — click column headers to sort",
    rownames = FALSE,
    options  = list(
      pageLength = 8,
      scrollX    = TRUE,
      dom        = "frtip",
      rowCallback = DT::JS(
        "function(row, data, index) {
           if (index % 2 === 0) {
             $(row).css({'background-color':'#8fa8c4','color':'#1a1a1a'});
           } else {
             $(row).css({'background-color':'#c8d8e8','color':'#1a1a1a'});
           }
         }"
      ),
      initComplete = DT::JS(
        "function(settings, json) {
           $(this.api().table().header()).css({'background-color':'#4a6080','color':'#ffffff'});
         }"
      )
    ),
    class = "compact hover"
  )
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

df |>
  select(driver_age, vehicle_age, vehicle_value, power_to_weight_ratio,
         total_premium, total_incurred, total_claims, total_exposure) |>
  summarise(across(everything(),
                   list(
                     Min    = ~min(., na.rm=TRUE),
                     Q1     = ~quantile(., 0.25, na.rm=TRUE),
                     Median = ~median(., na.rm=TRUE),
                     Mean   = ~mean(., na.rm=TRUE),
                     Q3     = ~quantile(., 0.75, na.rm=TRUE),
                     Max    = ~max(., na.rm=TRUE)
                   ))) |>
  pivot_longer(everything(), names_sep = "_(?=[^_]+$)", names_to = c("Variable","Stat")) |>
  pivot_wider(names_from = Stat, values_from = value) |>
  mutate(across(where(is.numeric), ~round(., 2))) |>
  DT::datatable(
    caption  = "Descriptive Statistics for Numerical Variables — click column headers to sort",
    rownames = FALSE,
    options  = list(pageLength = 8, scrollX = TRUE, dom = "frtip"),
    class = "compact stripe hover"
  )
```
:::

## Categorical Variables

### Justification of Technique

For categorical variables, **bar charts** provide the clearest representation of frequency distributions and proportional composition. Horizontal bars are preferred when category labels are long to avoid rotation. **Stacked** and **proportional** bar charts expose compositional relationships across grouping dimensions.

### Coverage Type and Policy Status

::: panel-tabset
#### The Plot

```{r cat-coverage}
#| echo: false
#| label: cat-coverage
#| fig-cap: "Coverage type distribution and policy status composition"

# ggiraph (Chap 03 §3.6): geom_col_interactive adds tooltip (hover text) and
# data_id (coordinated highlighting) — hovering one bar dims all others
p_pt <- df |>
  count(policy_type, sort = TRUE) |>
  mutate(policy_type = fct_reorder(policy_type, n),
         pct = n / sum(n),
         tooltip = paste0(policy_type,
                          "\nPolicies: ", scales::comma(n),
                          "\nShare: ", scales::percent(pct, accuracy = 0.1))) |>
  ggplot(aes(n, policy_type, fill = policy_type)) +
  geom_col_interactive(aes(tooltip = tooltip, data_id = policy_type),
                       colour = NA, alpha = 0.9) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            hjust = -0.1, size = 3.5) +
  scale_fill_manual(values = pal_cat5, guide = "none") +
  scale_x_continuous(labels = scales::comma,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Policies by Coverage Type",
       x = "Number of Policies", y = NULL)

p_bs_status <- df |>
  count(policy_status, business_type) |>
  group_by(business_type) |>
  mutate(pct = n / sum(n),
         tooltip = paste0(business_type, " — ", policy_status,
                          "\nCount: ", scales::comma(n),
                          "\nShare: ", scales::percent(pct, accuracy = 0.1))) |>
  ggplot(aes(business_type, n, fill = policy_status)) +
  geom_col_interactive(aes(tooltip = tooltip, data_id = policy_status),
                       position = "fill", colour = NA, width = 0.6) +
  scale_fill_manual(values = c(Active = "#4CAF50", Cancelled = "#F44336")) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Policy Status by Business Type",
       x = "Business Type", y = "Proportion", fill = "Status")

# girafe() wraps the patchwork into an interactive SVG widget
girafe(
  ggobj     = p_pt + p_bs_status,
  width_svg = 10, height_svg = 4,
  options   = list(
    opts_hover(css     = "fill-opacity:1; stroke:white; stroke-width:1.5px;"),
    opts_hover_inv(css = "opacity:0.25;"),
    opts_tooltip(css   = "background-color:#f5f5f5; color:#222;
                          padding:8px; border-radius:4px; font-size:12px;
                          border:1px solid #ccc;")
  )
)
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

p_pt <- df |>
  count(policy_type, sort = TRUE) |>
  mutate(policy_type = fct_reorder(policy_type, n),
         pct = n / sum(n)) |>
  ggplot(aes(n, policy_type, fill = policy_type)) +
  geom_col(colour = NA, alpha = 0.9) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            hjust = -0.1, size = 3.5) +
  scale_fill_manual(values = pal_cat5, guide = "none") +
  scale_x_continuous(labels = scales::comma,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Policies by Coverage Type",
       x = "Number of Policies", y = NULL)

p_bs_status <- df |>
  count(policy_status, business_type) |>
  ggplot(aes(business_type, n, fill = policy_status)) +
  geom_col(position = "fill", colour = NA, width = 0.6) +
  scale_fill_manual(values = c(Active = "#4CAF50", Cancelled = "#F44336")) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Policy Status by Business Type",
       x = "Business Type", y = "Proportion", fill = "Status")

p_pt + p_bs_status
```
:::

Comprehensive with excess (COMP_E) dominates at 54% of policies, reflecting consumer preference for broad coverage. Third-party liability with glass (TPG) is the second most common at \~25%, catering to budget-conscious customers seeking basic plus glass cover. The CC category — third-party liability combined with two or more additional coverages such as theft, fire, or total loss — represents a mid-tier segment between basic TP and full comprehensive. Notably, Cancelled policies are slightly more prevalent among New Business than Portfolio renewals, which is expected given higher lapse rates in the first policy year.

### Bonus Score and Payment Frequency

::: panel-tabset
#### The Plot

```{r cat-bonus}
#| echo: false
#| label: cat-bonus
#| fig-cap: "Bonus score distribution and payment frequency by coverage type"

p_bonus <- df |>
  count(bonus_score) |>
  filter(!is.na(bonus_score)) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(bonus_score, pct, fill = bonus_score)) +
  geom_col(colour = NA, width = 0.6) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            vjust = -0.5, size = 4) +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1.05)) +
  labs(title = "Bonus Score Distribution",
       subtitle = "96.5% hold Good status — positive portfolio quality signal",
       x = "Bonus Score", y = "Proportion")

p_pf <- df |>
  count(payment_frequency, policy_type) |>
  ggplot(aes(payment_frequency, n, fill = policy_type)) +
  geom_col(position = "fill", colour = NA, width = 0.6) +
  scale_fill_manual(values = pal_cat5) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Payment Frequency by Coverage",
       x = "Payment Frequency", y = "Proportion", fill = "Coverage")

p_bonus + p_pf
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

p_bonus <- df |>
  count(bonus_score) |>
  filter(!is.na(bonus_score)) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(bonus_score, pct, fill = bonus_score)) +
  geom_col(colour = NA, width = 0.6) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            vjust = -0.5, size = 4) +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1.05)) +
  labs(title = "Bonus Score Distribution",
       subtitle = "96.5% hold Good status — positive portfolio quality signal",
       x = "Bonus Score", y = "Proportion")

p_pf <- df |>
  count(payment_frequency, policy_type) |>
  ggplot(aes(payment_frequency, n, fill = policy_type)) +
  geom_col(position = "fill", colour = NA, width = 0.6) +
  scale_fill_manual(values = pal_cat5) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Payment Frequency by Coverage",
       x = "Payment Frequency", y = "Proportion", fill = "Coverage")

p_bonus + p_pf
```
:::

The bonus score reflects the policyholder's past claims experience: "Good (Favourable History)" indicates a favourable claims history, "Neutral" a moderate profile, and "Bad (Poor History)" denotes a poor claims record. The distribution is heavily concentrated toward Good (96.5%), limiting the discriminatory power of this variable for the majority of the portfolio. However, the 3.5% with Neutral or Bad scores are disproportionately important for risk segmentation, as demonstrated in the bivariate analysis. Annual payment is the dominant method across all coverage types, with quarterly payment more common in budget products (TP, TPG).

### Vehicle Fuel Type and Circulation Area

::: panel-tabset
#### The Plot

```{r cat-geo}
#| echo: false
#| label: cat-geo
#| fig-cap: "Fuel type and geographic area distributions"

p_fuel <- df |>
  filter(!is.na(fuel_type)) |>
  count(fuel_type) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(fuel_type, pct, fill = fuel_type)) +
  geom_col(colour = NA, width = 0.55) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            vjust = -0.5, size = 4.5) +
  scale_fill_manual(values = pal_fuel, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.65)) +
  labs(title = "Fuel Type Distribution",
       x = "Fuel Type", y = "Proportion")

p_area <- df |>
  count(circulation_area, municipality_type) |>
  ggplot(aes(circulation_area, n, fill = municipality_type)) +
  geom_col(position = "stack", colour = NA, width = 0.6) +
  scale_fill_manual(values = c("#1565C0","#42A5F5","#90CAF9")) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Area × Municipality Type",
       x = "Circulation Area", y = "Policies", fill = "Municipality")

p_fuel + p_area
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

p_fuel <- df |>
  filter(!is.na(fuel_type)) |>
  count(fuel_type) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(fuel_type, pct, fill = fuel_type)) +
  geom_col(colour = NA, width = 0.55) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            vjust = -0.5, size = 4.5) +
  scale_fill_manual(values = pal_fuel, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.65)) +
  labs(title = "Fuel Type Distribution",
       x = "Fuel Type", y = "Proportion")

p_area <- df |>
  count(circulation_area, municipality_type) |>
  ggplot(aes(circulation_area, n, fill = municipality_type)) +
  geom_col(position = "stack", colour = NA, width = 0.6) +
  scale_fill_manual(values = c("#1565C0","#42A5F5","#90CAF9")) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Area × Municipality Type",
       x = "Circulation Area", y = "Policies", fill = "Municipality")

p_fuel + p_area
```
:::

Gasoline vehicles (57%) slightly outnumber diesel (43%), broadly reflecting the Spanish vehicle fleet composition. Note that `fuel_type` records only these two categories — other fuel types (electric, hybrid) are not present in this dataset. The `circulation_area` variable classifies the predominant driving environment as Urban (U) or Rural (R). Continental urban policies dominate the geographic mix, consistent with Spain's population concentration in metropolitan areas. The `municipality_type` distinguishes Inland (I), Coastal (C), and Islands (IS) — an important consideration for territorial risk analysis since island policies show distinct risk profiles from mainland ones.

------------------------------------------------------------------------

# Bivariate Visual Analytics

Bivariate analysis investigates relationships between pairs of variables to identify factors influencing portfolio profitability and volatility. Each analysis is accompanied by a justification of the chosen visualisation technique and a discussion of statistical implications.

## Temporal Analysis: Year-Over-Year Trends

**Technique justification:** Line charts are optimal for displaying temporal trends in continuous metrics, allowing visual detection of trend direction and acceleration. Overlaid bar charts enable simultaneous comparison of absolute volumes and derived ratios. The static charts are converted to interactive `plotly` widgets via `ggplotly()` and linked into a `subplot()` with `shareX = TRUE`, so zooming or panning on the premium/incurred panel automatically adjusts the loss ratio panel — enabling coordinated exploration of the same time axis across both metrics.

::: panel-tabset
#### The Plot

```{r temporal-analysis}
#| echo: false
#| label: temporal-analysis
#| fig-cap: "Annual premium and claims trends with loss ratio evolution"

yearly <- df |>
  group_by(year) |>
  summarise(
    premium    = sum(total_premium, na.rm = TRUE),
    incurred   = sum(total_incurred, na.rm = TRUE),
    n_policies = n(),
    claim_freq = mean(has_claim, na.rm = TRUE),
    loss_ratio = incurred / premium,
    avg_premium = mean(total_premium, na.rm = TRUE),
    .groups = "drop"
  )

# plotly (Chap 03 §3.7): ggplotly() converts ggplot2 objects to interactive
# HTML widgets. subplot() with shareX = TRUE links x-axis zoom/pan across panels.
# hovermode = "x unified" shows all series for a year in one tooltip.
p_vol <- yearly |>
  pivot_longer(c(premium, incurred), names_to = "type", values_to = "amount") |>
  mutate(type = factor(type, levels = c("premium","incurred"),
                       labels = c("Premium Written","Claims Incurred")),
         hover_text = paste0(type, " (", year, ")<br>",
                             scales::dollar(amount/1e6, prefix="€", suffix="M", accuracy=0.1))) |>
  ggplot(aes(year, amount / 1e6, colour = type, group = type, text = hover_text)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 4) +
  scale_colour_manual(values = c("#1E88E5","#F44336")) +
  scale_y_continuous(labels = scales::dollar_format(prefix="€",suffix="M"),
                     limits = c(0, 65)) +
  labs(title = "Annual Premium Written vs Claims Incurred",
       x = "Year", y = "Amount (€M)", colour = NULL)

p_lr_yr <- yearly |>
  mutate(hover_text = paste0("Year: ", year,
                              "<br>Loss Ratio: ", scales::percent(loss_ratio, accuracy=0.1),
                              "<br>Policies: ", scales::comma(n_policies),
                              "<br>Claim Freq: ", scales::percent(claim_freq, accuracy=0.1))) |>
  ggplot(aes(year, loss_ratio, fill = loss_ratio, group = 1, text = hover_text)) +
  geom_col(alpha = 0.7, width = 0.5) +
  geom_point(colour = "white", size = 3) +
  scale_fill_gradient2(low = "#4CAF50", mid = "#FF9800", high = "#F44336",
                       midpoint = 0.70, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.85)) +
  labs(title = "Loss Ratio Trend", x = "Year", y = "Loss Ratio (%)")

subplot(
  ggplotly(p_vol, tooltip = "text"),
  ggplotly(p_lr_yr, tooltip = "text"),
  nrows = 2, heights = c(0.65, 0.35), shareX = TRUE
) |>
  layout(hovermode = "x unified")
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

yearly <- df |>
  group_by(year) |>
  summarise(
    premium    = sum(total_premium, na.rm = TRUE),
    incurred   = sum(total_incurred, na.rm = TRUE),
    n_policies = n(),
    claim_freq = mean(has_claim, na.rm = TRUE),
    loss_ratio = incurred / premium,
    avg_premium = mean(total_premium, na.rm = TRUE),
    .groups = "drop"
  )

p_vol <- yearly |>
  pivot_longer(c(premium, incurred), names_to = "type", values_to = "amount") |>
  mutate(type = factor(type, levels = c("premium","incurred"),
                       labels = c("Premium Written","Claims Incurred"))) |>
  ggplot(aes(year, amount / 1e6, colour = type, group = type)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 4) +
  geom_text(aes(label = scales::dollar(amount/1e6, prefix="€", suffix="M", accuracy=0.1)),
            vjust = -1.2, size = 3.5) +
  scale_colour_manual(values = c("#1E88E5","#F44336")) +
  scale_y_continuous(labels = scales::dollar_format(prefix="€",suffix="M"),
                     limits = c(0, 65)) +
  labs(title = "Annual Premium Written vs Claims Incurred",
       x = "Year", y = "Amount (€M)", colour = NULL)

p_lr_yr <- yearly |>
  ggplot(aes(year, loss_ratio, fill = loss_ratio, group = 1)) +
  geom_line(colour = "grey40", linewidth = 0.8) +
  geom_col(alpha = 0.7, width = 0.5) +
  geom_point(colour = "white", size = 3) +
  geom_text(aes(label = scales::percent(loss_ratio, accuracy = 0.1)),
            vjust = -0.7, size = 4, fontface = "bold") +
  scale_fill_gradient2(low = "#4CAF50", mid = "#FF9800", high = "#F44336",
                       midpoint = 0.70, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.85)) +
  labs(title = "Loss Ratio Trend",
       x = "Year", y = "Loss Ratio (%)")

p_vol / p_lr_yr + plot_layout(heights = c(2,1))
```
:::

**Statistical insight:** Premium written grew at a CAGR of approximately 65% over the two-year period (2022–2024), reflecting aggressive portfolio expansion. The loss ratio was stable at \~66% in both 2022 and 2023, then jumped sharply to \~75% in 2024 — a 9 percentage-point deterioration in a single year. This sudden 2024 spike, rather than a gradual drift, is more consistent with a large claims event, reserve strengthening, or a structural shift in the book than with general claims inflation. This warrants urgent investigation before any pricing action is taken.

## Driver Age vs Claim Frequency and Severity

**Technique justification:** Grouped bar charts with ordered categories (young to old) clearly communicate the monotonic relationship between age and risk. Separating frequency (incidence) from severity (cost given claim) follows the standard two-part actuarial model, which is more informative than aggregated loss ratios alone.

::: panel-tabset
#### The Plot

```{r driver-age-bivariate}
#| echo: false
#| label: driver-age-bivariate
#| fig-cap: "Claim frequency and average severity by driver age group"

age_summary <- df |>
  group_by(driver_age_group) |>
  summarise(
    n = n(),
    claim_freq   = mean(has_claim, na.rm = TRUE),
    avg_severity = mean(total_incurred[total_incurred > 0], na.rm = TRUE),
    loss_ratio   = sum(total_incurred, na.rm=TRUE) / sum(total_premium, na.rm=TRUE),
    .groups = "drop"
  )

# ggiraph coordinated views (Chap 03 §3.6.7): the same data_id on both charts
# means hovering an age group highlights it in BOTH frequency and severity plots
# simultaneously — revealing the frequency-severity trade-off at a glance.
age_summary <- age_summary |>
  mutate(
    tooltip_freq = paste0("Age Group: ", driver_age_group,
                          "\nClaim Frequency: ", scales::percent(claim_freq, accuracy=0.1),
                          "\nPolicies: ", scales::comma(n),
                          "\nLoss Ratio: ", scales::percent(loss_ratio, accuracy=0.1)),
    tooltip_sev  = paste0("Age Group: ", driver_age_group,
                          "\nAvg Severity: ", scales::dollar(avg_severity, prefix="€", accuracy=1),
                          "\nPolicies: ", scales::comma(n))
  )

p_freq_age <- age_summary |>
  ggplot(aes(driver_age_group, claim_freq, fill = claim_freq)) +
  geom_col_interactive(aes(tooltip = tooltip_freq, data_id = driver_age_group),
                       colour = NA, width = 0.7) +
  geom_text(aes(label = scales::percent(claim_freq, accuracy=0.1)),
            vjust = -0.4, size = 3.8, fontface = "bold") +
  scale_fill_gradient(low = "#42A5F5", high = "#F44336", guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.25)) +
  labs(title = "Claim Frequency by Driver Age Group",
       subtitle = "Monotonic decline from youngest (20.5%) to oldest (12.1%)",
       x = "Age Group", y = "Claim Frequency")

p_sev_age <- age_summary |>
  ggplot(aes(driver_age_group, avg_severity, fill = driver_age_group)) +
  geom_col_interactive(aes(tooltip = tooltip_sev, data_id = driver_age_group),
                       colour = NA, width = 0.7) +
  scale_fill_brewer(palette = "Blues", guide = "none") +
  geom_text(aes(label = scales::dollar(avg_severity, prefix = "€", accuracy = 1)),
            vjust = -0.4, size = 3.8) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "€"),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(title    = "Average Claim Severity by Driver Age Group",
       subtitle = "Severity does not mirror frequency — non-linear risk profile",
       x = "Age Group", y = "Avg Incurred per Claim (€)")

girafe(
  ggobj     = p_freq_age / p_sev_age,
  width_svg = 10, height_svg = 7,
  options   = list(
    opts_hover(css     = "fill-opacity:1; stroke:white; stroke-width:2px;"),
    opts_hover_inv(css = "opacity:0.2;"),
    opts_tooltip(css   = "background-color:#f5f5f5; color:#222;
                          padding:8px; border-radius:4px; font-size:12px;
                          border:1px solid #ccc;")
  )
)
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

age_summary <- df |>
  group_by(driver_age_group) |>
  summarise(
    n = n(),
    claim_freq   = mean(has_claim, na.rm = TRUE),
    avg_severity = mean(total_incurred[total_incurred > 0], na.rm = TRUE),
    loss_ratio   = sum(total_incurred, na.rm=TRUE) / sum(total_premium, na.rm=TRUE),
    .groups = "drop"
  )

p_freq_age <- age_summary |>
  ggplot(aes(driver_age_group, claim_freq, fill = claim_freq)) +
  geom_col(colour = NA, width = 0.7) +
  geom_text(aes(label = scales::percent(claim_freq, accuracy=0.1)),
            vjust = -0.4, size = 3.8, fontface = "bold") +
  scale_fill_gradient(low = "#42A5F5", high = "#F44336", guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.25)) +
  labs(title = "Claim Frequency by Driver Age Group",
       subtitle = "Monotonic decline from youngest (20.5%) to oldest (12.1%)",
       x = "Age Group", y = "Claim Frequency")

p_sev_age <- age_summary |>
  ggplot(aes(driver_age_group, avg_severity, fill = driver_age_group)) +
  geom_col(colour = NA, width = 0.7) +
  scale_fill_brewer(palette = "Blues", guide = "none") +
  geom_text(aes(label = scales::dollar(avg_severity, prefix = "€", accuracy = 1)),
            vjust = -0.4, size = 3.8) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "€"),
    expand = expansion(mult = c(0, 0.15))   # 15% headroom above tallest bar
  ) +
  labs(title    = "Average Claim Severity by Driver Age Group",
       subtitle = "Severity does not mirror frequency — non-linear risk profile",
       x = "Age Group", y = "Avg Incurred per Claim (€)")

p_freq_age / p_sev_age
```
:::

::: panel-tabset
#### The Plot

```{r driver-age-ridge}
#| echo: false
#| label: driver-age-ridge
#| fig-cap: "Premium distribution by driver age group (ridge plot)"

df |>
  filter(total_premium > 0, total_premium < 1500, !is.na(driver_age_group)) |>
  ggplot(aes(total_premium, driver_age_group, fill = driver_age_group)) +
  geom_density_ridges(alpha = 0.75, scale = 1.3, colour = NA) +
  scale_fill_brewer(palette = "RdYlBu", direction = -1, guide = "none") +
  scale_x_continuous(labels = scales::dollar_format(prefix="€")) +
  labs(title = "Premium Distribution by Driver Age Group",
       subtitle = "Ridge plot reveals premium spread and pricing differentiation by age",
       x = "Total Premium (€)", y = "Driver Age Group")
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

df |>
  filter(total_premium > 0, total_premium < 1500, !is.na(driver_age_group)) |>
  ggplot(aes(total_premium, driver_age_group, fill = driver_age_group)) +
  geom_density_ridges(alpha = 0.75, scale = 1.3, colour = NA) +
  scale_fill_brewer(palette = "RdYlBu", direction = -1, guide = "none") +
  scale_x_continuous(labels = scales::dollar_format(prefix="€")) +
  labs(title = "Premium Distribution by Driver Age Group",
       subtitle = "Ridge plot reveals premium spread and pricing differentiation by age",
       x = "Total Premium (€)", y = "Driver Age Group")
```
:::

**Statistical insight:** The monotonic decline in claim frequency from 20.5% (18–25) to 12.1% (66+) is statistically robust across the large sample. The ridge plot confirms that younger drivers (18–25) are charged substantially higher median premiums (\~€474) than all other age groups (\~€250–€262), validating that the pricing model incorporates age-related frequency risk. However, the severity analysis reveals an important counter-pattern: claim cost **increases** with driver age, with the 66+ cohort generating the highest average severity at \~€1,982 versus the lowest in the 26–35 group at \~€1,658. This means older drivers file claims less often but generate more expensive claims when they do — a frequency-severity trade-off that a simple loss ratio approach obscures. A two-part GLM (frequency × severity) is strongly recommended to correctly price both dimensions.

## Bonus Score vs Loss Ratio

**Technique justification:** Side-by-side bar charts with explicit loss ratio labels enable direct comparison across the three bonus-score categories. The ridge plot supplement shows premium adequacy — whether higher-risk segments are charged commensurately.

::: panel-tabset
#### The Plot

```{r bonus-bivariate}
#| echo: false
#| label: bonus-bivariate
#| fig-cap: "Loss ratio and claim frequency by bonus score"

bonus_summary <- df |>
  filter(!is.na(bonus_score)) |>
  group_by(bonus_score) |>
  summarise(
    n = n(),
    total_prem   = sum(total_premium, na.rm = TRUE),
    total_inc    = sum(total_incurred, na.rm = TRUE),
    claim_freq   = mean(has_claim, na.rm = TRUE),
    loss_ratio   = total_inc / total_prem,
    avg_premium  = mean(total_premium, na.rm = TRUE),
    .groups = "drop"
  )

p_bonus_lr <- bonus_summary |>
  ggplot(aes(bonus_score, loss_ratio, fill = bonus_score)) +
  geom_col(colour = NA, width = 0.6) +
  geom_text(aes(label = scales::percent(loss_ratio, accuracy = 0.1)),
            vjust = -0.5, size = 5, fontface = "bold") +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.9)) +
  labs(title = "Loss Ratio by Bonus Score",
       x = "Bonus Score", y = "Loss Ratio (%)")

p_bonus_cf <- bonus_summary |>
  ggplot(aes(bonus_score, claim_freq, fill = bonus_score)) +
  geom_col(colour = NA, width = 0.6) +
  geom_text(aes(label = scales::percent(claim_freq, accuracy = 0.1)),
            vjust = -0.5, size = 5, fontface = "bold") +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.4)) +
  labs(title = "Claim Frequency by Bonus Score",
       x = "Bonus Score", y = "Claim Frequency (%)")

p_bonus_lr + p_bonus_cf
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

bonus_summary <- df |>
  filter(!is.na(bonus_score)) |>
  group_by(bonus_score) |>
  summarise(
    n = n(),
    total_prem   = sum(total_premium, na.rm = TRUE),
    total_inc    = sum(total_incurred, na.rm = TRUE),
    claim_freq   = mean(has_claim, na.rm = TRUE),
    loss_ratio   = total_inc / total_prem,
    avg_premium  = mean(total_premium, na.rm = TRUE),
    .groups = "drop"
  )

p_bonus_lr <- bonus_summary |>
  ggplot(aes(bonus_score, loss_ratio, fill = bonus_score)) +
  geom_col(colour = NA, width = 0.6) +
  geom_text(aes(label = scales::percent(loss_ratio, accuracy = 0.1)),
            vjust = -0.5, size = 5, fontface = "bold") +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.9)) +
  labs(title = "Loss Ratio by Bonus Score",
       x = "Bonus Score", y = "Loss Ratio (%)")

p_bonus_cf <- bonus_summary |>
  ggplot(aes(bonus_score, claim_freq, fill = bonus_score)) +
  geom_col(colour = NA, width = 0.6) +
  geom_text(aes(label = scales::percent(claim_freq, accuracy = 0.1)),
            vjust = -0.5, size = 5, fontface = "bold") +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.4)) +
  labs(title = "Claim Frequency by Bonus Score",
       x = "Bonus Score", y = "Claim Frequency (%)")

p_bonus_lr + p_bonus_cf
```
:::

::: panel-tabset
#### The Plot

```{r bonus-premium-ridge}
#| echo: false
#| label: bonus-premium-ridge
#| fig-cap: "Premium distribution by bonus score — are higher-risk segments adequately priced?"

df |>
  filter(!is.na(bonus_score), total_premium > 0, total_premium < 1500) |>
  ggplot(aes(total_premium, bonus_score, fill = bonus_score)) +
  geom_density_ridges(alpha = 0.75, scale = 1.2, colour = NA) +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_x_continuous(labels = scales::dollar_format(prefix="€")) +
  labs(title = "Premium Distribution by Bonus Score",
       subtitle = "Bad-scored policyholders show slightly higher premium density at upper tail",
       x = "Annual Premium (€)", y = "Bonus Score")
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

df |>
  filter(!is.na(bonus_score), total_premium > 0, total_premium < 1500) |>
  ggplot(aes(total_premium, bonus_score, fill = bonus_score)) +
  geom_density_ridges(alpha = 0.75, scale = 1.2, colour = NA) +
  scale_fill_manual(values = pal_bonus, guide = "none") +
  scale_x_continuous(labels = scales::dollar_format(prefix="€")) +
  labs(title = "Premium Distribution by Bonus Score",
       subtitle = "Bad-scored policyholders show slightly higher premium density at upper tail",
       x = "Annual Premium (€)", y = "Bonus Score")
```
:::

**Statistical insight:** The "Bad (Poor History)" bonus score has a loss ratio of 79.0% versus 69.7% for "Good (Favourable History)" — a 9.3 percentage-point gap. Notably, the "Neutral" category has an even higher loss ratio at 81.4%, suggesting it captures policyholders in transition from Good to Bad rather than a genuinely moderate-risk group. The claim frequency gap between Bad (21.9%) and Good (14.2%) is substantial at 7.7 percentage points. The premium ridge plot confirms that Bad-scored policyholders pay higher premiums on average (mean €589 vs €290 for Good), but the persistent loss ratio gap indicates the current loading is still insufficient. A minimum 12–15% additional loading would be required to bring Bad scores to portfolio-average profitability.

## Coverage Type vs Profitability

**Technique justification:** Horizontal bar charts ordered by loss ratio enable rapid identification of the most and least profitable product lines. Violin plots (replacing the original scatter plot) show the full profit distribution per coverage type — including median, interquartile spread, and tail behaviour — which is more informative than a sample scatter for communicating volatility concentration to actuarial and management audiences.

::: panel-tabset
#### The Plot

```{r coverage-bivariate}
#| echo: false
#| label: coverage-bivariate
#| fig-cap: "Loss ratio and profitability metrics by coverage type"

cov_summary <- df |>
  group_by(policy_type) |>
  summarise(
    n = n(),
    avg_premium  = mean(total_premium, na.rm = TRUE),
    avg_incurred = mean(total_incurred, na.rm = TRUE),
    avg_profit   = mean(profit, na.rm = TRUE),
    loss_ratio   = sum(total_incurred, na.rm=TRUE) / sum(total_premium, na.rm=TRUE),
    claim_freq   = mean(has_claim, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(policy_type = fct_reorder(policy_type, loss_ratio))

p_cov_lr <- cov_summary |>
  ggplot(aes(loss_ratio, policy_type, fill = loss_ratio)) +
  geom_col(colour = NA) +
  geom_vline(xintercept = sum(df$total_incurred,na.rm=TRUE)/sum(df$total_premium,na.rm=TRUE),
             colour = "#FF9800", linetype = "dashed", linewidth = 1) +
  geom_text(aes(label = scales::percent(loss_ratio, accuracy = 0.1)),
            hjust = -0.1, size = 3.8, colour = "grey20") +
  scale_fill_gradient2(low = "#4CAF50", mid = "#FF9800", high = "#F44336",
                       midpoint = 0.70, guide = "none") +
  scale_x_continuous(labels = scales::percent,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Loss Ratio by Coverage Type",
       subtitle = "Dashed line = portfolio average (70.3%)",
       x = "Loss Ratio", y = NULL)

p_profit_box <- df |>
  filter(profit > -5000, profit < 2000) |>
  ggplot(aes(policy_type, profit, fill = policy_type)) +
  geom_violin(alpha = 0.7, colour = NA, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_hline(yintercept = 0, colour = "#F44336", linetype = "dashed") +
  scale_fill_manual(values = pal_cat5, guide = "none") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "€")) +
  labs(title = "Profit Distribution by Coverage Type",
       subtitle = "Violin plot shows spread and skewness of per-policy profits",
       x = NULL, y = "Profit per Policy (€)")

p_cov_lr + p_profit_box + plot_layout(widths = c(1, 1.2))
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

cov_summary <- df |>
  group_by(policy_type) |>
  summarise(
    n = n(),
    avg_premium  = mean(total_premium, na.rm = TRUE),
    avg_incurred = mean(total_incurred, na.rm = TRUE),
    avg_profit   = mean(profit, na.rm = TRUE),
    loss_ratio   = sum(total_incurred, na.rm=TRUE) / sum(total_premium, na.rm=TRUE),
    claim_freq   = mean(has_claim, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(policy_type = fct_reorder(policy_type, loss_ratio))

p_cov_lr <- cov_summary |>
  ggplot(aes(loss_ratio, policy_type, fill = loss_ratio)) +
  geom_col(colour = NA) +
  geom_vline(xintercept = sum(df$total_incurred,na.rm=TRUE)/sum(df$total_premium,na.rm=TRUE),
             colour = "#FF9800", linetype = "dashed", linewidth = 1) +
  geom_text(aes(label = scales::percent(loss_ratio, accuracy=0.1)),
            hjust = -0.1, size = 3.8) +
  scale_fill_gradient2(low = "#4CAF50", mid = "#FF9800", high = "#F44336",
                       midpoint = 0.70, guide = "none") +
  scale_x_continuous(labels = scales::percent,
                     expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Loss Ratio by Coverage Type",
       subtitle = "Dashed line = portfolio average (70.3%)",
       x = "Loss Ratio", y = NULL)

p_profit_box <- df |>
  filter(profit > -5000, profit < 2000) |>
  ggplot(aes(policy_type, profit, fill = policy_type)) +
  geom_violin(alpha = 0.7, colour = NA, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_hline(yintercept = 0, colour = "#F44336", linetype = "dashed") +
  scale_fill_manual(values = pal_cat5, guide = "none") +
  scale_y_continuous(labels = scales::dollar_format(prefix="€")) +
  coord_flip() +
  labs(title = "Profit Distribution by Coverage Type",
       subtitle = "Violin plot shows spread and skewness of per-policy profits",
       x = NULL, y = "Profit per Policy (€)")

p_cov_lr / p_profit_box
```
:::

**Statistical insight:** Third Party achieves the best loss ratio (\~54.3%) owing to its limited coverage scope, while Comprehensive (No Excess) (COMP_N) is the most loss-making at \~93.9% — far above the portfolio average of 70.3%. This is actuarially expected: COMP_N policyholders bear no excess on claims, removing the cost-sharing incentive that suppresses small claims. Third Party+Multi-Cover (CC) performs better than expected at \~63.4% despite its multi-coverage structure, likely due to selective underwriting of this mid-tier segment. Comprehensive (With Excess) (COMP_E) sits close to the portfolio average at \~73.8%, reflecting that the excess loading partially compensates for the broader coverage. The violin plots confirm that COMP_N and COMP_E carry the widest downside profit distributions, representing the greatest per-policy volatility.

## Vehicle Value vs Loss Ratio

**Technique justification:** Quintile-based bar charts reduce noise from the continuous vehicle value variable while maintaining ordinal structure. This technique, known as "equal-frequency binning", is standard in actuarial GLM rating factor visualisation. The charts are implemented as native `plot_ly()` widgets arranged in a `subplot()`, enabling rich hover tooltips (value range, loss ratio, claim frequency, avg premium, policy count) without cluttering the chart with on-bar text labels.

::: panel-tabset
#### The Plot

```{r vehicle-value-bivariate}
#| echo: false
#| label: vehicle-value-bivariate
#| fig-cap: "Loss ratio and claim frequency by vehicle value quintile — hover bars for exact values and details"

veh_quintiles <- df |>
  filter(!is.na(vehicle_value), vehicle_value > 0) |>
  group_by(veh_value_q) |>
  summarise(
    min_val    = min(vehicle_value),
    max_val    = max(vehicle_value),
    n          = n(),
    loss_ratio = sum(total_incurred, na.rm=TRUE) / sum(total_premium, na.rm=TRUE),
    claim_freq = mean(has_claim, na.rm = TRUE),
    avg_premium = mean(total_premium, na.rm=TRUE),
    .groups = "drop"
  ) |>
  mutate(
    label = paste0("Q", veh_value_q, "\n€",
                   scales::comma(round(min_val/1000)),"k-",
                   scales::comma(round(max_val/1000)),"k")
  )

# plotly native plot_ly() (Chap 03 §3.7.1): more control than ggplotly().
# subplot() creates a coordinated two-panel view with shared x-axis.
veh_quintiles <- veh_quintiles |>
  mutate(
    value_range = paste0("€", scales::comma(round(min_val/1000)), "k – €",
                         scales::comma(round(max_val/1000)), "k"),
    hover_lr    = paste0("Quintile: Q", veh_value_q,
                         "<br>Value range: ", value_range,
                         "<br>Loss Ratio: ", scales::percent(loss_ratio, accuracy=0.1),
                         "<br>Policies: ", scales::comma(n),
                         "<br>Avg Premium: €", round(avg_premium, 0)),
    hover_cf    = paste0("Quintile: Q", veh_value_q,
                         "<br>Value range: ", value_range,
                         "<br>Claim Frequency: ", scales::percent(claim_freq, accuracy=0.1),
                         "<br>Policies: ", scales::comma(n))
  )

# textposition = "none" removes bar text labels — values visible on hover instead
p_lr <- plot_ly(veh_quintiles, x = ~label, y = ~loss_ratio,
                type = "bar", name = "Loss Ratio",
                marker = list(color = ~loss_ratio,
                              colorscale = list(c(0,"#4CAF50"),c(0.5,"#FF9800"),c(1,"#F44336")),
                              showscale = FALSE),
                text = ~hover_lr, hoverinfo = "text",
                textposition = "none") |>
  layout(yaxis = list(title = "Loss Ratio", tickformat = ".1%"),
         xaxis = list(title = "Vehicle Value Quintile"),
         title = "U-shaped: mid-value vehicles most profitable")

p_cf <- plot_ly(veh_quintiles, x = ~label, y = ~claim_freq,
                type = "bar", name = "Claim Frequency",
                marker = list(color = ~claim_freq,
                              colorscale = list(c(0,"#90CAF9"),c(1,"#1565C0")),
                              showscale = FALSE),
                text = ~hover_cf, hoverinfo = "text",
                textposition = "none") |>
  layout(yaxis = list(title = "Claim Frequency", tickformat = ".1%"),
         xaxis = list(title = "Vehicle Value Quintile"),
         title = "Claim frequency rises monotonically Q1 → Q5")

subplot(p_lr, p_cf, nrows = 1, shareX = TRUE, titleX = TRUE) |>
  layout(showlegend = FALSE)
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

veh_quintiles <- df |>
  filter(!is.na(vehicle_value), vehicle_value > 0) |>
  group_by(veh_value_q) |>
  summarise(
    min_val    = min(vehicle_value),
    max_val    = max(vehicle_value),
    n          = n(),
    loss_ratio = sum(total_incurred, na.rm=TRUE) / sum(total_premium, na.rm=TRUE),
    claim_freq = mean(has_claim, na.rm = TRUE),
    avg_premium = mean(total_premium, na.rm=TRUE),
    .groups = "drop"
  ) |>
  mutate(
    label = paste0("Q", veh_value_q, "\n€",
                   scales::comma(round(min_val/1000)),"k-",
                   scales::comma(round(max_val/1000)),"k")
  )

p_vval_lr <- veh_quintiles |>
  ggplot(aes(label, loss_ratio, fill = loss_ratio)) +
  geom_col(colour = NA, width = 0.7) +
  geom_text(aes(label = scales::percent(loss_ratio, accuracy=0.1)),
            vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_gradient2(low = "#4CAF50", mid = "#FF9800", high = "#F44336",
                       midpoint = 0.70, guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.88)) +
  labs(title = "Loss Ratio by Vehicle Value Quintile",
       subtitle = "Q1 = lowest value vehicles; Q5 = highest value vehicles",
       x = "Vehicle Value Quintile", y = "Loss Ratio (%)")

p_vval_cf <- veh_quintiles |>
  ggplot(aes(label, claim_freq, fill = claim_freq)) +
  geom_col(colour = NA, width = 0.7) +
  geom_text(aes(label = scales::percent(claim_freq, accuracy=0.1)),
            vjust = -0.5, size = 4) +
  scale_fill_gradient(low = "#90CAF9", high = "#1565C0", guide = "none") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.22)) +
  labs(title = "Claim Frequency by Vehicle Value Quintile",
       x = "Vehicle Value Quintile", y = "Claim Frequency (%)")

p_vval_lr + p_vval_cf
```
:::

**Statistical insight:** Hover over each bar to see the full detail (value range, loss ratio, claim frequency, policy count, avg premium). Contrary to the initial hypothesis, vehicle value does not produce a simple monotonic relationship with loss ratio. Loss ratios across quintiles are Q1: 72.3%, Q2: 69.3%, Q3: 69.0%, Q4: 67.2%, Q5: 73.3% — a U-shaped pattern where mid-value vehicles (Q3–Q4) are actually the most profitable. The top quintile (Q5, vehicles over \~€34,000) has a slightly elevated loss ratio at 73.3%, primarily driven by severity on expensive repairs. Claim **frequency**, however, does rise monotonically from 10.9% (Q1) to 17.9% (Q5), confirming higher-value vehicles are more frequently claimed. The vehicle value loading in theft and property damage premiums (correlation r ≈ 0.33–0.41) appears to partially but not fully compensate for the elevated frequency at the top end.

## Urban vs Rural Dimension

**Technique justification:** Two complementary interactive charts are used. First, a `ggplotly()` line chart enables hover-inspection of each year-area data point with full metric detail (loss ratio, claim frequency, avg premium, policy count). Second, a `plot_ly()` bar chart with `frame = ~year` provides a year-selection slider, allowing the reader to directly compare the urban-rural claim frequency gap across age groups for any chosen year — a more analytical interaction than a looping animation.

::: panel-tabset
#### The Plot

```{r urban-rural}
#| echo: false
#| label: urban-rural
#| fig-cap: "Urban vs Rural loss ratio, claim frequency, and premium trends"

area_year <- df |>
  group_by(circulation_area, year) |>
  summarise(
    n = n(),
    loss_ratio = sum(total_incurred,na.rm=TRUE)/sum(total_premium,na.rm=TRUE),
    claim_freq = mean(has_claim, na.rm=TRUE),
    avg_premium = mean(total_premium, na.rm=TRUE),
    .groups = "drop"
  )

# plotly (Chap 03 §3.7): ggplotly converts the trend line to an interactive widget
area_year <- area_year |>
  mutate(hover = paste0(circulation_area, " (", year, ")",
                        "<br>Loss Ratio: ", scales::percent(loss_ratio, accuracy=0.1),
                        "<br>Claim Freq: ", scales::percent(claim_freq, accuracy=0.1),
                        "<br>Avg Premium: €", round(avg_premium, 0),
                        "<br>Policies: ", scales::comma(n)))

p_lr_area_yr <- area_year |>
  ggplot(aes(year, loss_ratio, colour=circulation_area, group=circulation_area,
             text=hover)) +
  geom_line(linewidth=1.4) +
  geom_point(size=4) +
  scale_colour_manual(values=pal_area) +
  scale_y_continuous(labels=scales::percent, limits=c(0.55,0.82)) +
  # Push x-axis title down via margin to prevent overlap with the plotly legend
  labs(title="Loss Ratio Trend: Urban vs Rural (hover for full details)",
       x="Year", y="Loss Ratio", colour="Area") +
  theme(axis.title.x = element_text(margin = margin(t = 30)))

# Move legend to top-right corner inside the plot area to free x-axis space
ggplotly(p_lr_area_yr, tooltip = "text") |>
  layout(legend = list(x = 0.85, y = 0.15,
                       bgcolor = "rgba(255,255,255,0.7)",
                       bordercolor = "#cccccc", borderwidth = 1))
```

```{r urban-rural-year-filter}
#| echo: false
#| label: urban-rural-year-filter
#| fig-cap: "Claim frequency by age group and area — use the slider to select year"

# plotly frame slider: plot_ly() with frame = ~year renders a Play button and
# step-through slider. The user can click any year directly or press Play to
# animate. animation_opts() controls transition speed; layout() positions the
# slider label. Note: barmode must be set in layout(), not plot_ly(), when
# frame is used — otherwise plotly cannot resolve the grouped bar structure.
age_area_year <- df |>
  group_by(circulation_area, driver_age_group, year) |>
  summarise(claim_freq = mean(has_claim, na.rm=TRUE), .groups="drop") |>
  mutate(
    hover_text = paste0(circulation_area, " — Age ", driver_age_group,
                        "<br>Year: ", year,
                        "<br>Claim Frequency: ", scales::percent(claim_freq, accuracy=0.1))
  )

plot_ly(age_area_year,
        x         = ~driver_age_group,
        y         = ~claim_freq,
        color     = ~circulation_area,
        colors    = c(Urban = "#FF9800", Rural = "#1E88E5"),
        frame     = ~year,
        type      = "bar",
        text      = ~hover_text,
        hoverinfo = "text",
        textposition = "none") |>
  layout(
    title   = "Claim Frequency by Age Group and Area",
    barmode = "group",          # must be in layout() when frame is used
    xaxis   = list(title = "Driver Age Group"),
    yaxis   = list(title = "Claim Frequency", tickformat = ".1%",
                   range = c(0, 0.28)),
    legend  = list(title = list(text = "Area"))
  ) |>
  animation_opts(
    frame      = 800,
    transition = 400,
    easing     = "cubic-in-out",
    redraw     = FALSE
  )
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

area_year <- df |>
  group_by(circulation_area, year) |>
  summarise(
    n = n(),
    loss_ratio = sum(total_incurred,na.rm=TRUE)/sum(total_premium,na.rm=TRUE),
    claim_freq = mean(has_claim, na.rm=TRUE),
    avg_premium = mean(total_premium, na.rm=TRUE),
    .groups = "drop"
  )

# Chart 1: ggplotly interactive loss ratio trend
area_year <- area_year |>
  mutate(hover = paste0(circulation_area, " (", year, ")",
                        "<br>Loss Ratio: ", scales::percent(loss_ratio, accuracy=0.1),
                        "<br>Claim Freq: ", scales::percent(claim_freq, accuracy=0.1),
                        "<br>Avg Premium: \u20ac", round(avg_premium, 0),
                        "<br>Policies: ", scales::comma(n)))

p_lr_area_yr <- area_year |>
  ggplot(aes(year, loss_ratio, colour=circulation_area, group=circulation_area,
             text=hover)) +
  geom_line(linewidth=1.4) +
  geom_point(size=4) +
  scale_colour_manual(values=pal_area) +
  scale_y_continuous(labels=scales::percent, limits=c(0.55,0.82)) +
  labs(title="Loss Ratio Trend: Urban vs Rural (hover for full details)",
       x="Year", y="Loss Ratio", colour="Area") +
  theme(axis.title.x = element_text(margin = margin(t = 30)))

ggplotly(p_lr_area_yr, tooltip = "text") |>
  layout(legend = list(x = 0.85, y = 0.15,
                       bgcolor = "rgba(255,255,255,0.7)",
                       bordercolor = "#cccccc", borderwidth = 1))

# Chart 2: plotly year-slider bar chart
age_area_year <- df |>
  group_by(circulation_area, driver_age_group, year) |>
  summarise(claim_freq = mean(has_claim, na.rm=TRUE), .groups="drop") |>
  mutate(hover_text = paste0(circulation_area, " — Age ", driver_age_group,
                              "<br>Year: ", year,
                              "<br>Claim Frequency: ",
                              scales::percent(claim_freq, accuracy=0.1)))

plot_ly(age_area_year, x=~driver_age_group, y=~claim_freq,
        color=~circulation_area, colors=c(Urban="#FF9800", Rural="#1E88E5"),
        frame=~year, type="bar",
        text=~hover_text, hoverinfo="text", textposition="none") |>
  layout(title="Claim Frequency by Age Group and Area",
         barmode = "group",
         xaxis=list(title="Driver Age Group"),
         yaxis=list(title="Claim Frequency", tickformat=".1%", range=c(0,0.28)),
         legend=list(title=list(text="Area"))) |>
  animation_opts(frame=800, transition=400, easing="cubic-in-out", redraw=FALSE)
```
:::

**Statistical insight:** Urban policyholders have a higher overall loss ratio (71.4% vs 68.9% rural), a gap of 2.5 percentage points. The interactive trend chart above allows hovering each data point for the full breakdown (loss ratio, claim frequency, average premium, policy count) per year and area. Crucially, the year-by-year data reveals that in 2022 rural policies (70.5%) were actually less profitable than urban (61.8%), with positions stabilising in 2023 before diverging sharply in 2024 (urban: 76.3%, rural: 72.6%). Use the year slider in the chart below to compare the age-group claim frequency patterns across 2022, 2023 and 2024 — the urban disadvantage emerges clearly only from 2023 onward, and is most pronounced for the 18-25 cohort (22.2% urban vs 18.9% rural in 2024). This urban-age interaction supports a two-dimensional rating factor structure rather than treating geography and age as independent additive factors.

## Correlation Analysis

**Technique justification:** The Pearson correlation heatmap provides a compact overview of linear relationships across multiple numerical variables simultaneously. This is a standard first-pass tool in feature engineering for actuarial models, identifying redundant variables (high multicollinearity) and informative predictors (moderate correlations with claims outcomes).

::: panel-tabset
#### The Plot

```{r correlation}
#| echo: false
#| label: correlation
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Pearson correlation matrix for premium components and risk variables"

cor_vars <- df |>
  select(
    Total_Premium   = total_premium,
    Liab_Prem       = liability_premium,
    Property_Prem   = property_damage_premium,
    Theft_Prem      = theft_premium,
    Glass_Prem      = glass_premium,
    Total_Incurred  = total_incurred,
    Total_Claims    = total_claims,
    Driver_Age      = driver_age,
    Vehicle_Age     = vehicle_age,
    Veh_Value       = vehicle_value,
    Pwr_Weight      = power_to_weight_ratio,
    Exposure        = total_exposure
  )

corr_mat <- cor(cor_vars, use = "complete.obs", method = "pearson")

ggcorrplot(
  corr_mat,
  method    = "square",
  type      = "lower",
  lab       = TRUE,
  lab_size  = 4.5,
  colors    = c("#D32F2F","#FAFAFA","#1565C0"),
  outline.color = "white",
  tl.cex    = 12,
  tl.col    = "grey20"
) +
  labs(title = "Pearson Correlation Matrix",
       subtitle = "Premium components, claims outcomes, and risk characteristics") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### The Code

```{r}
#| eval: false
#| code-fold: false

cor_vars <- df |>
  select(
    Total_Premium   = total_premium,
    Liab_Prem       = liability_premium,
    Property_Prem   = property_damage_premium,
    Theft_Prem      = theft_premium,
    Glass_Prem      = glass_premium,
    Total_Incurred  = total_incurred,
    Total_Claims    = total_claims,
    Driver_Age      = driver_age,
    Vehicle_Age     = vehicle_age,
    Veh_Value       = vehicle_value,
    Pwr_Weight      = power_to_weight_ratio,
    Exposure        = total_exposure
  )

corr_mat <- cor(cor_vars, use = "complete.obs", method = "pearson")

ggcorrplot(
  corr_mat,
  method    = "square",
  type      = "lower",
  lab       = TRUE,
  lab_size  = 3,
  colors    = c("#D32F2F","#FAFAFA","#1565C0"),
  outline.color = "white",
  tl.cex    = 10,
  tl.col    = "grey20"
) +
  labs(title = "Pearson Correlation Matrix",
       subtitle = "Premium components, claims outcomes, and risk characteristics") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
:::

**Statistical insight:** Premium components are moderately to strongly correlated with total premium (liability r = 0.68, property damage r = 0.75, theft r = 0.49, glass r = 0.52), reflecting their derivation from a common pricing model. Total incurred claims show a **very weak** positive correlation with premium (r ≈ 0.07) — considerably weaker than commonly assumed — underscoring the highly stochastic nature of insurance outcomes where premium is a prospective expectation and incurred is a realised outcome dominated by rare large losses. Vehicle value is positively correlated with total premium (r ≈ 0.32) and with theft and property damage premiums (r ≈ 0.33–0.41), validating actuarially sound rating factor application, though the relationship is weaker than the original text implied. Driver age has a very weak negative correlation with claim counts (r ≈ −0.03), consistent with the frequency-age relationship identified in Section 5.2 but indicating that age alone explains minimal variance in claims — multivariate modelling is essential.

------------------------------------------------------------------------

# Summary of Visual Analytics Techniques

```{r technique-summary}
#| label: technique-summary
#| echo: false

techniques <- tibble(
  Technique         = c(
    "Histogram + Density Overlay",
    "Ridge Plot (ggridges)",
    "Bar Chart (sorted)",
    "Stacked/Proportional Bar",
    "Interactive Line Chart (plotly)",
    "Interactive Grouped Bar (ggiraph)",
    "Violin Plot",
    "Correlation Heatmap (ggcorrplot)",
    "Interactive Bar Chart (plotly)",
    "Interactive DataTable (DT)",
    "Year-Slider Bar Chart (plotly)"
  ),
  `Package / Function` = c(
    "geom_histogram + geom_density",
    "geom_density_ridges (ggridges)",
    "geom_col + fct_reorder",
    "geom_col (position=fill)",
    "ggplotly() + subplot() (plotly)",
    "geom_col_interactive + girafe() (ggiraph)",
    "geom_violin",
    "ggcorrplot()",
    "plot_ly() + subplot() (plotly)",
    "DT::datatable() (DT)",
    "plot_ly(frame=~year) (plotly)"
  ),
  `Variable Types`  = c(
    "Continuous",
    "Continuous × Categorical",
    "Categorical",
    "Categorical × Categorical",
    "Continuous × Temporal",
    "Categorical × Continuous",
    "Categorical × Continuous",
    "Continuous × Continuous (multi)",
    "Continuous × Continuous",
    "Mixed (summary table)",
    "Categorical × Continuous × Temporal"
  ),
  `Analytical Purpose` = c(
    "Reveal distributional shape and skewness",
    "Compare premium/loss distributions across groups",
    "Rank category frequencies; spot imbalances",
    "Reveal compositional differences across groups",
    "Trend exploration with hover detail and linked zoom",
    "Coordinated hover highlights age group across both frequency & severity charts",
    "Show full profit distribution including tails",
    "Identify collinearity and informative predictors",
    "Quintile risk comparison with rich hover tooltips",
    "Sortable, searchable summary statistics table",
    "Year-by-year claim frequency comparison via user-controlled slider"
  )
)

library(htmltools)
ts_css <- '
<style>
  #tech-tbl th {
    background-color: #4a6080; color: #ffffff; border-bottom: 2px solid #2d4a6a;
  }
  #tech-tbl tr:nth-child(odd)  td { background-color: #8fa8c4; color: #1a1a1a; }
  #tech-tbl tr:nth-child(even) td { background-color: #c8d8e8; color: #1a1a1a; }
  #tech-tbl tr:hover td { background-color: #7a9ab8; color: #1a1a1a; }
</style>
'
ts_tbl <- kbl(techniques,
              caption = "Visual Analytics Techniques Applied in This Investigation",
              table.attr = 'id="tech-tbl"') |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE)
ts_clean <- gsub('style="([^"]*)(background-color:[^;]+;?)([^"]*)"',
                  'style="\\1\\3"', as.character(ts_tbl))
HTML(paste0(ts_css, ts_clean))
```

------------------------------------------------------------------------

# Key Findings and Analytical Conclusions

## Profitability

1.  **Overall loss ratio of 70.3%** was stable at \~66% in 2022–2023, then **spiked to 75% in 2024** — a 9 pp single-year deterioration requiring urgent root cause investigation before pricing action.
2.  **Comprehensive (No Excess) (COMP_N)** is the least profitable product line (\~93.9% loss ratio); the absence of any excess removes the policyholder's cost-sharing incentive.
3.  **Third Party** policies generate the best underwriting margins (\~54.3%) but represent only 1.6% of the portfolio.
4.  **Vehicle value** shows a U-shaped relationship with loss ratio — mid-value vehicles (Q3–Q4) are the most profitable; claim **frequency** rises monotonically from 10.9% (Q1) to 17.9% (Q5).

## Risk Segmentation Factors

5.  **Bonus score** discriminates clearly: both Bad (79.0%) and Neutral (81.4%) policyholders significantly exceed Good (69.7%), with Neutral actually the worst-performing segment.
6.  **Driver age** shows a monotonic inverse relationship with claim frequency (20.5% → 12.1%) but a **positive** relationship with severity (€1,658 for 26–35 up to €1,982 for 66+) — the two dimensions move in opposite directions and must be priced separately.
7.  **Urban location** adds \~2.5 pp to the overall loss ratio, with a sharp 2024 divergence (urban 76.3% vs rural 72.6%).

## Volatility

8.  Claim severity is **highly right-skewed** — a small proportion of large claims drive total incurred costs disproportionately.
9.  The correlation between premium and incurred (r ≈ 0.07) is extremely weak, confirming individual policy outcomes are largely unpredictable and portfolio diversification is essential.

## Recommended Actions

```{r}
#| echo: false
rec_df <- tibble(
  Priority = c("High","High","High","Medium","Medium","Low"),
  Action = c(
    "Investigate 2024 loss ratio spike before pricing changes",
    "Review COMP_N pricing urgently — consider introducing a mandatory minimum excess",
    "Increase bonus-malus loading for Neutral and Bad scores",
    "Introduce age-specific severity loading for 66+ drivers",
    "Strengthen urban territorial factors for 2025 pricing",
    "Monitor vehicle value frequency trend"
  ),
  Rationale = c(
    "9 pp single-year jump — needs root cause analysis",
    "93.9% loss ratio; no cost-sharing incentive",
    "Neutral (81.4%) actually worse than Bad (79.0%)",
    "Highest severity group despite lowest frequency",
    "Urban-rural gap widening; 2024 urban LR at 76.3%",
    "Monotonic frequency rise Q1→Q5 not yet fully reflected in pricing"
  )
)

rec_css <- '
<style>
  #rec-tbl th {
    background-color: #4a6080; color: #ffffff; border-bottom: 2px solid #2d4a6a;
  }
  #rec-tbl tr:nth-child(odd)  td { background-color: #8fa8c4; color: #1a1a1a; }
  #rec-tbl tr:nth-child(even) td { background-color: #c8d8e8; color: #1a1a1a; }
  #rec-tbl tr:hover td { background-color: #7a9ab8; color: #1a1a1a; }
</style>
'

rec_tbl <- kbl(rec_df, booktabs = TRUE, table.attr = 'id="rec-tbl"') |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = TRUE) |>
  column_spec(1, width = "8em",  bold = TRUE) |>
  column_spec(2, width = "28em") |>
  column_spec(3, width = "24em")

rec_clean <- gsub('style="([^"]*)(background-color:[^;]+;?)([^"]*)"',
                   'style="\\1\\3"', as.character(rec_tbl))
HTML(paste0(rec_css, rec_clean))
```
